### 2021-04-15 tidymodels_analysis.Rmd
r markdown report of results

---
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, comment = "")

library(tidyverse)
library(nortest)
library(FSA)
library(tidymodels)
library(factoextra)
library(plotly)
library(usemodels)
library(ranger)
library(vip)

```

## Dysregulation of hemoglobin and oxygenation in cerebral malaria {.tabset}

```{r data}

# read in data
df_hbox <- read.csv("Data/98 first visit records with cerebral oxygenation stats_trimmed.csv") %>% as_tibble()

# get rid of "Other"
df_hbox <- df_hbox %>% 
  filter(Status != "Other") %>% 
  mutate(Status = factor(Status, levels = c("HV", "UM", "CM")))

# identify duplicated rows (TM0003, TM2001)
df_hbox <- df_hbox %>% filter(duplicated(Subject.ID..NIAID.) == FALSE)

# Convert age to numeric
df_hbox <- df_hbox %>% 
  mutate(Age = gsub(" Years, ", "_", Age)) %>% 
  mutate(Age = gsub(" Months ", "", Age)) %>% 
  separate(Age, into = c("Years", "Months"), sep = "_") %>% 
  mutate(Years = as.numeric(Years)) %>% 
  mutate(Months = as.numeric(Months)/12) %>% 
  mutate(Age = Years + Months) %>% 
  select(-c(Years, Months))

# Convert sex to factor
df_hbox <- df_hbox %>% mutate(Sex = as.factor(Sex))

# Trim for variables of interest (VOI)
df_hbox_trim <- df_hbox %>% 
  dplyr::select(c(Subject.ID..NIAID., Status, Glucose, Hematocrit, Lactate, Temperature,
                  Arginine.umol.L, Haptoglobin..mg.dl., Hemoglobin..uM., Whole.Blood.Nitrite, 
                  cerebral.hbdeox, cerebral.hbdeox.alpha, cerebral.hbox, cerebral.hbox.alpha,
                  cerebral.sat, cerebral.sat.alpha, cerebral.thc, cerebral.thc.alpha, cerebral.thc.norm.std)) %>%

  dplyr::rename("subject_id" = "Subject.ID..NIAID.", "Arginine (umol/L)" = "Arginine.umol.L", "Haptoglobin (mg/dL)" = "Haptoglobin..mg.dl.", 
                "Hemoglobin (uM)" = "Hemoglobin..uM.", "WB_Nitrite" = "Whole.Blood.Nitrite", "DeoxyHb" = "cerebral.hbdeox", 
                "DeoxyHb_alpha" = "cerebral.hbdeox.alpha", "OxyHb" = "cerebral.hbox", "OxyHb_alpha" = "cerebral.hbox.alpha", 
                "O2sat" = "cerebral.sat", "O2sat_alpha" = "cerebral.sat.alpha", "THC" = "cerebral.thc", 
                "THC_alpha" = "cerebral.thc.alpha", "THC_sd" = "cerebral.thc.norm.std")

# impute missing data with median by group
f_impute <- function(x, na.rm = TRUE) (replace(x, is.na(x), median(x, na.rm = na.rm)))

df_hbox_impute <- df_hbox_trim %>% 
  group_by(Status) %>% 
  mutate_at(3:ncol(.), f_impute) %>% 
  ungroup()

# pivot long
df_hbox_long <- df_hbox_trim %>% pivot_longer(cols = 3:ncol(.), names_to = "measurement", values_to = "value")

```

### Exploratory analysis

#### Histogram to explore normality of data
```{r plot1, fig.width = 15, fig.height = 8}

df_hbox_long %>%

  ggplot(aes(x = value)) +
  geom_histogram(bins = 20) +
  facet_grid(Status ~ measurement, scales = "free") +

  theme(strip.text.y = element_text(size = 15),
        strip.text.x = element_text(size = 11),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 0.9))

```

#### Plots of all the variables by status
```{r plot2, fig.width = 15, fig.height = 15}

df_hbox_long %>%

  ggplot(aes(x = Status, y = value)) +
  geom_violin(aes(color = Status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = Status), size = 0.2, width = 0.5) +
  facet_wrap(~measurement, scales = "free_y") +

  theme_bw() +
  theme(legend.position = "none")  +

  theme(strip.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        axis.title.y = element_blank(),
        axis.text = element_text(size = 15))

```

### PCA

#### Can we cluster patients based on available variables?

```{r pca_data}

# convert tibble to df
df_hbox_impute_df <- df_hbox_impute %>% select(-c(subject_id, Status)) %>% as.data.frame(df_hbox_impute) 
rownames(df_hbox_impute_df) <- df_hbox_impute$subject_id

# compute PCA
my_pca <- prcomp(df_hbox_impute_df, scale = TRUE)

# define groups
groups <- as.factor(df_hbox_impute$Status)

# Eigenvalues
pca_eig_val <- get_eigenvalue(my_pca)

# Results for Variables
pca_res_var <- get_pca_var(my_pca)

# Results for individuals
pca_res_ind <- get_pca_ind(my_pca)

```

#### 2D PCA
```{r pca_2d, fig.width = 10, fig.height = 10}

p_pca <- fviz_pca_ind(my_pca,
                      col.ind = groups, # color by groups
                      palette = c("#00AFBB",  "#FC4E07", "#7CAE00"),
                      addEllipses = TRUE, # Concentration ellipses
                      ellipse.type = "confidence",
                      legend.title = "Groups",
                      repel = TRUE
)

p_pca$layers[[4]] <- NULL # get rid of text labels

p_pca

```

#### Biplot
```{r pca_biplot, fig.width = 12, fig.height = 12}

# Biplot of individuals and variables
p_biplot <- fviz_pca_biplot(my_pca, repel = TRUE,
                            col.var = "black", # Variables color
                            col.ind = groups  # Individuals color
)

p_biplot$layers[[3]] <- NULL
p_biplot$layers[[2]] <- NULL

p_biplot

```

#### Determine biggest contributors to each PC
```{r pca_contrib}

df_pca_contrib <- pca_res_var$contrib %>% 
  as_tibble(rownames = "variable") %>% 
  arrange(-Dim.1, -Dim.2)

df_contrib <- tibble(Dim.1 = paste0(df_pca_contrib$variable, 
                                    sep = ": ", 
                                    round(df_pca_contrib$Dim.1, 2)),
                     Dim.2 = paste0(arrange(df_pca_contrib, -Dim.2)$variable, 
                                    sep = ": ", 
                                    round(arrange(df_pca_contrib, -Dim.2)$Dim.2, 2)),
                     Dim.3 = paste0(arrange(df_pca_contrib, -Dim.3)$variable, 
                                    sep = ": ", 
                                    round(arrange(df_pca_contrib, -Dim.3)$Dim.3, 2)))

df_contrib %>% as.data.frame()

```

#### 3D PCA

```{r pca_3d}

df_pca <- pca_res_ind$coord %>% 
  as_tibble(rownames = "subject") %>% 
  mutate(Status = df_hbox_impute$Status)

plot_ly(x = df_pca$Dim.1, y = df_pca$Dim.2, z = df_pca$Dim.3, 
        type = "scatter3d", mode = "markers", color = df_pca$Status)

```

### Random forest predictive modeling: all data

#### Build, tune, and finalize model
- Random forest model
- bootstrap resampling (w/ replacement)
- BoxCox transform all predictors, remove variables with zero variance, kNN imputation
- tune for optimal ROC AUC and accuracy

```{r model1}

# Data
hbox_df <- df_hbox_trim %>% select(-subject_id)

# Build a model

  # split data set
  set.seed(123)
  hbox_split <- initial_split(hbox_df, strata = Status)
  hbox_train <- training(hbox_split)
  hbox_test <- testing(hbox_split)

  # resamples (not a lot of data, need to use cross validation)
  set.seed(234)
  # vfold_cv(hbox_train, strata = Status) # test too small
  hbox_folds <- bootstraps(hbox_train, strata = Status)

  # Scaffolding for setting up common types of models
  # use_ranger(Status ~ ., data = hbox_df)

  # Copy usemodels code...

  # Create recipe
  ranger_recipe <- 
    recipe(formula = Status ~ ., data = hbox_df) %>%

    # BoxCox transformation to normality
    step_BoxCox(all_predictors(), -all_outcomes()) %>%

    # remove variables with non-zero variance
    step_nzv(all_predictors(), -all_outcomes()) %>%

    # impute missing data using K-nearest neighbors
    step_knnimpute(all_predictors(), -all_outcomes())

  # Model specifications (set for tuning)
  ranger_spec <- 
    rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
    set_mode("classification") %>% 
    set_engine("ranger")

  # Create workflow
  ranger_workflow <- 
    workflow() %>% 
    add_recipe(ranger_recipe) %>% 
    add_model(ranger_spec)

# Tune model

  # Initial tune on bootstraps
  set.seed(84374)
  ranger_tune <-
    tune_grid(ranger_workflow, 
              resamples = hbox_folds, 
              grid = 11)

# Finalize model

  # finalize workflow
  final_rf <- ranger_workflow %>% 
    finalize_workflow(select_best(ranger_tune, metric = "accuracy"))

  final_rs_res <- final_rf %>% 
    fit_resamples(resamples = hbox_folds,
                  metrics = metric_set(roc_auc, accuracy, sens, spec),
                  control = control_resamples(save_pred = TRUE))

  hbox_fit <- last_fit(final_rf, hbox_split) # fitting to training data, evaluating on testing

```

#### Results

```{r model1_res}

# computed on test set
collect_metrics(hbox_fit) %>% select(.metric, .estimate) %>% as.data.frame()

# collect predictions on test set
df_pred <- collect_predictions(hbox_fit) %>%
  select(c(Status, .pred_HV, .pred_UM, .pred_CM, .pred_class))

df_pred %>% as_data_frame()

df_pred %>% filter(Status != .pred_class) %>% as.data.frame() # only got one wrong

```

```{r model1_plot_res}

conf_mat_resampled(final_rs_res) %>%

  ggplot(aes(x = Truth, y = Freq, fill = Prediction)) + 
  geom_col(position = "dodge") +
  ggtitle("Confusion matrix") +
  theme_bw()

```

#### Determine variable importance
```{r model1_imp}

# new specification for ranger model
imp_spec <- ranger_spec %>% 
  finalize_model(select_best(ranger_tune, metric = "accuracy")) %>% 
  set_engine("ranger", importance = "permutation")

# plot importance 
workflow() %>% 
  add_recipe(ranger_recipe) %>% 
  add_model(imp_spec) %>% 
  fit(hbox_train) %>% 
  pull_workflow_fit() %>% 
  vip(aesthetics = list(alpha = 0.8, fill = "midnightblue")) +

  theme_bw() +
  ggtitle("Variable importance")

```

### Random forest predictive modeling: NIRS data only

#### Build, tune, and finalize model
Same code, but built model only using variables measured using NIRS

```{r model2}

# Data
hbox_df2 <- df_hbox_impute %>% 
  select(c(Status, DeoxyHb, DeoxyHb_alpha, OxyHb, OxyHb_alpha, O2sat, O2sat_alpha, THC, THC_alpha, THC_sd))

# Build a model

  # split data set
  set.seed(456)
                hbox_split2 <- initial_split(hbox_df2, strata = Status)
                hbox_train2 <- training(hbox_split2)
                hbox_test2 <- testing(hbox_split2)

  # resamples (not a lot of data, need to use cross validation)
  set.seed(789)
  hbox_folds2 <- bootstraps(hbox_train2, strata = Status)

  # Scaffolding for setting up common types of models
  # use_ranger(Status ~ ., data = hbox_df2)

  # Copy usemodels code...

  # Create recipe
  ranger_recipe2 <- 
    recipe(formula = Status ~ ., data = hbox_df2) %>%

    # BoxCox transformation to normality
    step_BoxCox(all_predictors(), -all_outcomes()) %>%

    # remove variables with non-zero variance
    step_nzv(all_predictors(), -all_outcomes()) %>%

    # impute missing data using K-nearest neighbors
    step_knnimpute(all_predictors(), -all_outcomes())

  # Model specifications (set for tuning)
  ranger_spec2 <- 
    rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
    set_mode("classification") %>% 
    set_engine("ranger")

  # Create workflow
  ranger_workflow2 <- 
    workflow() %>% 
    add_recipe(ranger_recipe2) %>% 
    add_model(ranger_spec2)

# Tune model

  # Initial tune on bootstraps
  set.seed(101112)
  ranger_tune2 <-
    tune_grid(ranger_workflow2, 
              resamples = hbox_folds2, 
              grid = 11)

# Finalize model

  # finalize workflow
  final_rf2 <- ranger_workflow2 %>% 
    finalize_workflow(select_best(ranger_tune2, metric = "accuracy"))

  final_rs_res2 <- final_rf2 %>% 
    fit_resamples(resamples = hbox_folds2,
                  metrics = metric_set(roc_auc, accuracy, sens, spec),
                  control = control_resamples(save_pred = TRUE))

  hbox_fit2 <- last_fit(final_rf2, hbox_split2) # fitting to training data, evaluating on testing

```

#### Results

```{r model2_res}

# computed on test set
collect_metrics(final_rs_res2) %>% select(.metric, mean, std_err, n) %>% as.data.frame()

# collect predictions on test set
df_pred2 <- collect_predictions(hbox_fit2) %>%
  select(c(Status, .pred_HV, .pred_UM, .pred_CM, .pred_class))

df_pred2 %>% as_data_frame()

df_pred2 %>% filter(Status != .pred_class) %>% as.data.frame() # only got one wrong

```

```{r model2_plot_res}

conf_mat_resampled(final_rs_res2) %>%

  ggplot(aes(x = Truth, y = Freq, fill = Prediction)) + 
  geom_col(position = "dodge") +
  ggtitle("Confusion matrix") +
  theme_bw()

```

#### Determine variable importance
```{r model2_imp}

# new specification for ranger model
imp_spec2 <- ranger_spec2 %>% 
  finalize_model(select_best(ranger_tune2, metric = "accuracy")) %>% 
  set_engine("ranger", importance = "permutation")

# plot importance 
workflow() %>% 
  add_recipe(ranger_recipe2) %>% 
  add_model(imp_spec2) %>% 
  fit(hbox_train2) %>% 
  pull_workflow_fit() %>% 
  vip(aesthetics = list(alpha = 0.8, fill = "midnightblue")) +

  theme_bw() +
  ggtitle("Variable importance")

```