### 2021-06-29 figures.Rmd
updated figures

---
title: "Dysregulation of Cerebral Hemoglobin Content in Malawian Children with Malaria"
author: "Rachel Smith"
date: "`r Sys.Date()`"
output: html_document
---

<style type="text/css">
  body{
  font-family: Arial;
  font-size: 12pt;
}
</style>

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "", fig.height = 8, fig.width = 10)

library(tidyverse)
library(janitor)
library(FSA)
library(factoextra)
library(ggrepel)
library(ggpubr)
library(tidymodels)
library(vip)
library(scales)
library(reshape)
library(extrafont)
library(grDevices)

load("glm_model.Rdata")

font_import()

```

```{r data}

# read in data
df_hbox <- read.csv("Data/98 first visit records with cerebral oxygenation stats_trimmed.csv") %>% as_tibble()

# get rid of "Other"
df_hbox <- df_hbox %>% 
  filter(Status != "Other") %>% 
  mutate(Status = factor(Status, levels = c("HV", "UM", "CM")))

# identify duplicated rows (TM0003, TM2001)
df_hbox <- df_hbox %>% filter(duplicated(Subject.ID..NIAID.) == FALSE)

# Convert age to numeric
df_hbox <- df_hbox %>% 
  mutate(Age = gsub(" Years, ", "_", Age)) %>% 
  mutate(Age = gsub(" Months ", "", Age)) %>% 
  separate(Age, into = c("Years", "Months"), sep = "_") %>% 
  mutate(Years = as.numeric(Years)) %>% 
  mutate(Months = as.numeric(Months)/12) %>% 
  mutate(Age = Years + Months) %>% 
  dplyr::select(-c(Years, Months))

# Convert sex to factor
df_hbox <- df_hbox %>% mutate(Sex = as.factor(Sex))

# Trim for variables of interest (VOI)
df_hbox_trim <- df_hbox %>% 
  dplyr::select(c(Subject.ID..NIAID., Status, Glucose, Hematocrit, Lactate, Temperature,
                  Arginine.umol.L, Haptoglobin..mg.dl., Hemoglobin..uM., 
                  BP.Diastolic, BP.Systolic, O2.Sat, RR, HR,

                  cerebral.hbdeox, cerebral.hbdeox.alpha, cerebral.hbox, cerebral.hbox.alpha,
                  cerebral.sat, cerebral.sat.alpha, cerebral.thc, cerebral.thc.alpha)) %>%

  dplyr::rename("subject_id" = "Subject.ID..NIAID.", 
                "Arginine (umol/L)" = "Arginine.umol.L", "Haptoglobin (mg/dL)" = "Haptoglobin..mg.dl.", 
                "Hemoglobin (uM)" = "Hemoglobin..uM.", "Lactate (mmol/L)" = "Lactate", "Hematocrit (%)" = "Hematocrit",
                "Systolic BP" = "BP.Systolic", "Diastolic BP" = "BP.Diastolic", "O2 sat" = "O2.Sat", 
                "Respiration rate" = "RR", "Heart rate" = "HR",

                "[Hb]" = "cerebral.hbdeox", "[Hb]α" = "cerebral.hbdeox.alpha", 
                "[HbO2]" = "cerebral.hbox", "[HbO2]α" = "cerebral.hbox.alpha",  
                "HbO2%" = "cerebral.sat", "HbO2% α" = "cerebral.sat.alpha", 
                "[HbT]" = "cerebral.thc", "[HbT]α" = "cerebral.thc.alpha")

# Remove patients labeled as "UM" that instead meet the criteria for severe malaria

  # remove UM patient with lactate > 5 
  df_hbox_trim <- df_hbox_trim %>% dplyr::filter(!(Status == "UM" & `Lactate (mmol/L)` > 5))

  # All UM patients have hematocrit > 15%
  # df_hbox_trim %>% dplyr::filter(Status == "UM"  & `Hematocrit (%)` <= 15)

# impute missing data with median by group
f_impute <- function(x, na.rm = TRUE) (replace(x, is.na(x), median(x, na.rm = na.rm)))

df_hbox_impute <- df_hbox_trim %>% 
  group_by(Status) %>% 
  mutate_at(3:ncol(.), f_impute) %>% 
  ungroup()

# pivot long
df_hbox_long <- df_hbox_trim %>% pivot_longer(cols = 3:ncol(.), names_to = "variable", values_to = "value")

```

### Table 1: Summary of clinical data by patient group

```{r table1}

# number of subjects 
df_status <- df_hbox_impute %>% 
  count(Status) %>% 
  pivot_wider(1:2, names_from = "Status", values_from = "n") %>% 
  mutate(variable = "Number of subjects") %>% 
  dplyr::select(variable, everything()) %>% 
  mutate_if(is.numeric, as.character)

# Sex
df_sex <- df_hbox %>% 
  group_by(Status) %>% 
  count(Sex) %>% 
  filter(Sex != "") %>% 
  pivot_wider(names_from = "Sex", values_from = "n") %>% 
  mutate(perc_fem = round(Female/(Female + Male)*100, 2)) %>%

  pivot_wider(c(1,4), names_from = "Status", values_from = "perc_fem") %>% 
  mutate(variable = "Sex (% female)") %>% 
  dplyr::select(variable, everything()) %>% 
  mutate_if(is.numeric, as.character)

# other numeric variables

df_clinical_res <- df_hbox %>%

  dplyr::filter(Subject.ID..NIAID. != "TM1018") %>%

  group_by(Status) %>% 
  summarise(`Age (years)` = median(Age, na.rm = TRUE), age_q1 = quantile(Age, 0.25, na.rm = TRUE), age_q3 = quantile(Age, 0.75, na.rm = TRUE),

            `Temperature (°C)` = median(Temperature, na.rm = TRUE), temp_q1 = quantile(Temperature, 0.25, na.rm = TRUE), temp_q3 = quantile(Temperature, 0.75, na.rm = TRUE),
            `Heart rate` = median(HR, na.rm = TRUE), hr_q1 = quantile(HR, 0.25, na.rm = TRUE), hr_q3 = quantile(HR, 0.75, na.rm = TRUE),
            `Systolic BP` = median(BP.Systolic, na.rm = TRUE), sys_q1 = quantile(BP.Systolic, 0.25, na.rm = TRUE), sys_q3 = quantile(BP.Systolic, 0.75, na.rm = TRUE),
            `Diastolic BP` = median(BP.Diastolic, na.rm = TRUE), dia_q1 = quantile(BP.Diastolic, 0.25, na.rm = TRUE), dia_q3 = quantile(BP.Diastolic, 0.75, na.rm = TRUE),
            `Respiration rate` = median(RR, na.rm = TRUE), rr_q1 = quantile(RR, 0.25, na.rm = TRUE), rr_q3 = quantile(RR, 0.75, na.rm = TRUE),
            `O~2~ saturation` = median(O2.Sat, na.rm = TRUE), o2_q1 = quantile(O2.Sat, 0.25, na.rm = TRUE), o2_q3 = quantile(O2.Sat, 0.75, na.rm = TRUE),

            `Hematocrit (%)` = median(Hematocrit, na.rm = TRUE), hem_q1 = quantile(Hematocrit, 0.25, na.rm = TRUE), hem_q3 = quantile(Hematocrit, 0.75, na.rm = TRUE),
            `Lactate (mmol/L)` = median(Lactate, na.rm = TRUE), lac_q1 = quantile(Lactate, 0.25, na.rm = TRUE), lac_q3 = quantile(Lactate, 0.75, na.rm = TRUE),
            `Arginine (μmol/L)` = median(Arginine.umol.L, na.rm = TRUE), arg_q1 = quantile(Arginine.umol.L, 0.25, na.rm = TRUE), arg_q3 = quantile(Arginine.umol.L, 0.75, na.rm = TRUE)) %>% 
  mutate_if(is.numeric, round, 1) %>%

  mutate(`Age (years)` = paste0(`Age (years)`, " (", age_q1, ", ", age_q3, ")"),

         `Temperature (°C)` = paste0(`Temperature (°C)`, " (", temp_q1, ", ", temp_q3, ")"),
         `Heart rate` = paste0(`Heart rate`, " (", hr_q1, ", ", hr_q3, ")"),
         `Systolic BP` = paste0(`Systolic BP`, " (", sys_q1, ", ", sys_q3, ")"),
         `Diastolic BP` = paste0(`Diastolic BP`, " (", dia_q1, ", ", dia_q3, ")"),
         `Respiration rate` = paste0(`Respiration rate`, " (", rr_q1, ", ", rr_q3, ")"),
         `O~2~ saturation` = paste0(`O~2~ saturation`, " (", o2_q1, ", ", o2_q3, ")"),

         `Hematocrit (%)` = paste0(`Hematocrit (%)`, " (", hem_q1, ", ", hem_q3, ")"),
         `Lactate (mmol/L)` = paste0(`Lactate (mmol/L)`, " (", lac_q1, ", ", lac_q3, ")"),
         `Arginine (μmol/L)` = paste0(`Arginine (μmol/L)`, " (", arg_q1, ", ", arg_q3, ")")) %>% 
  dplyr::select(-c(age_q1, age_q3, 
                   temp_q1, temp_q3, hr_q1, hr_q3, sys_q1, sys_q3, dia_q1, dia_q3, rr_q1, rr_q3, o2_q1, o2_q3, 
                   hem_q1, hem_q3, lac_q1, lac_q3, arg_q1, arg_q3)) %>%

  unique() %>% 
  t() %>% 
  row_to_names(row_number = 1) %>% 
  as_tibble(rownames = "variable") %>% 
  add_row(df_status, .before = 1) %>% 
  add_row(df_sex, .before = 2) %>%

  dplyr::rename(" " = "variable",
                "Healthy volunteers" = "HV",
                "Uncomplicated malaria patients" = "UM",
                "Cerebral malaria patients" = "CM")

  # remove lactate data for HV
  df_clinical_res[11,]$`Healthy volunteers` <- "--"

  # create kable df
  df_clinical_res %>% knitr::kable(align = rep('c', 4))

```
**Table 1.** A summary of subject information and hematological data. Group patient data presented as median (quartile 1, quartile 3). Note: Sex not recorded for 3 CM patients; Lactate data removed for healthy volunteers due to inconsistent data.

### Figure 1: Visual summary of NIRS variables

```{r fig1}

df_hbox_voi <- df_hbox_impute %>% 
  dplyr::select(c(Status, `[HbO2]`, `[HbO2]α`, `[Hb]`, `[Hb]α`, `HbO2%`, `HbO2% α`, `[HbT]`, `[HbT]α`)) %>% 
  dplyr::filter(`[Hb]α` < 3)

df_hbox_voi_long <- df_hbox_voi %>% 
  pivot_longer(2:9, names_to = "variable", values_to = "value") %>%
  mutate(variable = factor(variable, levels = c("[HbO2]", "[HbO2]α",   
                                                 "[Hb]", "[Hb]α", 
                                                "HbO2%", "HbO2% α", 
                                                "[HbT]", "[HbT]α")))

my_col <- c("CM" = "#00BFC4", "HV" = "#7CAE00", "UM" = "#F8766D")

df_hbox_voi_long %>% dplyr::filter(variable %in% c("[HbO2]", "[Hb]", "HbO2%", "[HbT]")) %>%

  ggplot(aes(x = Status, y = value)) +
  geom_violin(aes(color = Status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = Status), size = 0.2, width = 0.5) +
  facet_wrap(~variable, scales = "free_y", nrow = 2, ncol = 2) +

  scale_color_manual(values = my_col) +
  theme_bw() +
  theme(legend.position = "none")  +

  theme(text = element_text(family = "Arial"),
        strip.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        axis.title.y = element_blank(),
        axis.text = element_text(size = 15))

```

```{r fig1.pvals, eval = FALSE}

# [HbT]
kruskal.test(`[HbT]` ~ Status, data = df_hbox_voi)
dunnTest(`[HbT]` ~ Status, data = df_hbox_voi, method = "bh")

# HbO2%
kruskal.test(`HbO2%` ~ Status, data = df_hbox_voi)

```

**Figure 1.** Shown are the variables measured by Oxiplex, plotted by patient status [HbT] levels were different between all three groups (Kruskal-Wallis adjusted p-value = 3.25x10^-6^; Dunn Test post-hoc adjusted p-values: CM vs HV = 1.66x10^-6^, CM vs UM = .032, HV vs UM = .033). Levels of cerebral O~2~ saturation did not differ among groups (Kruskal-Wallis adjusted p-value = .07).

### Figure 2: Visual summary of α values associated with NIRS variables

```{r fig2}

df_hbox_voi <- df_hbox_impute %>% 
  dplyr::select(c(Status, `[HbO2]`, `[HbO2]α`, `[Hb]`, `[Hb]α`, `HbO2%`, `HbO2% α`, `[HbT]`, `[HbT]α`)) %>% 
  dplyr::filter(`[Hb]α` < 3)

df_hbox_voi_long <- df_hbox_voi %>% 
  pivot_longer(2:9, names_to = "variable", values_to = "value") %>%
  mutate(variable = factor(variable, levels = c("[HbO2]", "[HbO2]α",   
                                                 "[Hb]", "[Hb]α", 
                                                "HbO2%", "HbO2% α", 
                                                "[HbT]", "[HbT]α")))

my_col <- c("CM" = "#00BFC4", "HV" = "#7CAE00", "UM" = "#F8766D")

df_hbox_voi_long %>% dplyr::filter(variable %in% c("[HbO2]α", "[Hb]α", "HbO2% α", "[HbT]α")) %>%

  ggplot(aes(x = Status, y = value)) +
  geom_violin(aes(color = Status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = Status), size = 0.2, width = 0.5) +

  geom_hline(yintercept = 0.4, alpha = 0.4, lty = 2) +
  geom_hline(yintercept = 0.6, alpha = 0.4, lty = 2) +
  geom_hline(yintercept = 1.0, alpha = 0.4, lty = 2) +
  geom_hline(yintercept = 1.5, alpha = 0.4, lty = 2) +

  facet_wrap(~variable, scales = "free_y", nrow = 2, ncol = 2) +

  ylim(0, 1.5) +
  scale_color_manual(values = my_col) +
  theme_bw() +

  theme(text = element_text(family = "Arial"),
        strip.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        legend.position = "none")

```

```{r fig2.pvals, eval = FALSE}

# [HbT]α
kruskal.test(`[HbT]α` ~ Status, data = df_hbox_voi)
dunnTest(`[HbT]α` ~ Status, data = df_hbox_voi, method = "bh")

```

**Figure 2** Shown are the alpha values associated with variables measured by the Oxiplex, plotted by patient status. Note: One outlier was removed from Deoxy Hb α (3.52).

### Figure 3: Logistic regression model on CM vs UM

#### A. PCA
```{r fig3a}
# include variables of interest ([HbT], [HbT] alpha, and hematocrit), remove healthy volunteers from analysis
df_hbox_log_voi <- df_hbox_impute %>% 
  dplyr::select(c(2,4:5,19:22)) %>% 
  dplyr::filter(Status != "HV")

# convert tibble to df
df_hbox_voi_df <- df_hbox_log_voi %>% 
  dplyr::select(-Status) %>% 
  as.data.frame() 
rownames(df_hbox_voi_df) <- df_hbox_impute %>% dplyr::filter(Status != "HV") %>% .$subject_id

# compute PCA
my_pca <- prcomp(df_hbox_voi_df, scale = TRUE)

# define groups
groups <- as.factor(df_hbox_log_voi$Status)

# Eigenvalues
pca_eig_val <- get_eigenvalue(my_pca)

# two data frames for plotting
df_vars <- my_pca$rotation %>% as_tibble() %>% mutate(variable = rownames(my_pca$rotation))
df_pnts <- my_pca$x %>% as_tibble() %>% mutate(group = groups)

# plot

  # plot patients
  ggplot() +
  geom_point(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = group), size = 3) +
  stat_conf_ellipse(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = group, fill = group),
                    level = 0.95, npoint = 100, bary = TRUE, alpha = 0.1, geom = "polygon") +
  stat_mean(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = group, shape = group),
            na.rm = FALSE) +

  # plot variables
  geom_point(data = df_vars, aes(x = PC1*5, y = PC2*5)) +
  geom_text_repel(data = df_vars, aes(x = PC1*5, y = PC2*5, label = variable, size = 10), show.legend = FALSE) +
  geom_segment(data = df_vars, aes(x = 0, y = 0, xend = PC1*5, yend = PC2*5), arrow = arrow()) +

  # draw lines
  geom_vline(xintercept = 0, lty = 2, color = "black") +
  geom_hline(yintercept = 0, lty = 2, color = "black") +

  # design
  xlab(paste0("PC1: ", round(pca_eig_val$variance.percent[1], 2), "% of variance")) +
  ylab(paste0("PC2: ", round(pca_eig_val$variance.percent[2], 2), "% of variance")) +

  theme_bw() +
  theme(text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))

```

#### B. Confusion matrix

```{r fig3b_prep, eval = FALSE}

# select [HbT], [HbT] alpha, and hematocrit
  df_hbox_model <- df_hbox_impute %>% 
    dplyr::select(c(2,4:5,19:22)) %>% 
    mutate(`1/10 * [HbT]α` = `[HbT]α`/10, `[HbT] (nM)` = `[HbT]`/1000) %>% 
    dplyr::select(-c(`[HbT]α`, `[HbT]`)) %>% 
    mutate(Status = as.factor(ifelse(Status == "CM", 1, 0))) %>%

    dplyr::select(-c(`HbO2%`, `HbO2% α`, `Lactate (mmol/L)`))

# resample
set.seed(234)
hbox_boot <- bootstraps(df_hbox_model, times = 1000)

# recipe
hbox_rec <- recipe(Status ~ ., data = df_hbox_model) %>% 
  step_BoxCox(all_predictors()) %>% 
  step_nzv(all_predictors())

# create workflow
hbox_wf <- workflow() %>% add_recipe(hbox_rec)

# logistic regression model specifications
glm_spec <- logistic_reg() %>% set_engine("glm")

# create model on 1000 different resamples
glm_res <- hbox_wf %>% 
  add_model(glm_spec) %>%
  fit_resamples(
    resamples = hbox_boot,
    metrics = metric_set(roc_auc, accuracy, sensitivity, specificity),
    control = tune::control_resamples(save_pred = TRUE, verbose = TRUE)
  )

# select best model to be used for evaluation
glm_best <- glm_res %>% select_best("roc_auc")

# finalize model
hbox_final <- hbox_wf %>% 
  add_model(glm_spec) %>% 
  finalize_workflow(glm_best) %>%
  fit(df_hbox_model) %>%
  pull_workflow_fit()

# save objects so model doesn't rerun every time we knit
save(glm_res, hbox_final, file = "glm_model.Rdata")

```

```{r fig3b_plot}

final_res_conf_mat <- glm_res %>% 
  conf_mat_resampled() %>% 
  mutate(Prediction = ifelse(Prediction == 1, "CM", "UM")) %>% 
  mutate(Truth = ifelse(Truth == 1, "CM", "UM"))

final_res_conf_mat %>% 
  group_by(Truth) %>% 
  summarise(total_true = sum(Freq)) %>% 
  right_join(final_res_conf_mat) %>% 
  mutate(`Percent occurence` = round(Freq/total_true*100, 2)) %>%

  # plot
  ggplot(aes(x = Truth, y = Prediction, fill = `Percent occurence`, label = `Percent occurence`)) +
  geom_tile() +
  geom_text(size = 6) +
  scale_fill_gradient2(low = "#4575b4", high = "#d73027") +

  theme_bw() +
  theme(text = element_text(family = "Arial"),
        axis.title = element_text(size = 25),
        axis.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 15))

```

**Figure 3.** Logistic regression model discriminating CM vs UM patients using cerebral tissue hemoglobin concentration, cerebral tissue hemoglobin concentration alpha, and hematocrit as the regressors **(A)** PCA of patients with CM and UM as determined by NIRS variables. Color indicates patient status, ellipses show median and 95% confidence interval on the cluster. **(B)** Confusion matrix representing the results of the logistic regression model averaged across 1000 bootstrap resamples. Percent occurrence indicates percentage times that the event (shown by the x- and y-axes) occurred.

### Table 2. Logistic regression model evaluation

#### A. Model metrics
```{r table2a}

collect_metrics(glm_res) %>% 
  dplyr::select(.metric, mean, std_err) %>% 
  mutate(mean = mean*100, std_err = std_err*100) %>% 
  mutate(Metric = c("Accuracy", "ROC AUC", "Sensitivity", "Specificity")) %>% 
  dplyr::select(c("Metric", "mean", "std_err")) %>%

  dplyr::rename("Mean" = "mean", "Standard error" = "std_err") %>% 
  mutate_if(is.numeric, round, 2) %>% 
  as.data.frame() %>%

  knitr::kable(align = rep('c', 4))

```

#### B. Coefficients
```{r table2b}

hbox_final %>% 
  tidy(exponentiate = TRUE) %>% 
  mutate(p.value = as.character(p.value)) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  mutate(p.value = scientific(p.value, digits = 3)) %>% 
  dplyr::rename("Term" = "term", "Odds ratio" = "estimate", "Standard error" = "std.error", 
                "Statistic" = "statistic", "p-value" = "p.value") %>%

  as.data.frame() %>%
  knitr::kable(align = rep('c', 5))

```

**Table 2.** Evaluation of the best logistic regression model select from 1000.

### Figure 4: Relationships among [HbT], [HbT] alpha, and hematocrit
```{r fig4}

df_hbox_impute %>% filter(Status != "HV") %>%

  ggplot(aes(x = `[HbT]`, y = `Hematocrit (%)`)) +
  geom_point(aes(color = Status, size = `[HbT]α`), alpha = 0.7) +

  xlab("[HbT] (μM)") +

  theme_bw() +
  theme(text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))

```

**Figure 4.**

### Supplementary figure 1: Spearman correlations of potential regressors
```{r fig_s1, fig.show = "hold", out.width = "50%", fig.height = 12, fig.width = 15}

# calculate spearman correlations in each group

f_calc_cor <- function(df, status) {

  df %>%
    filter(Status == status) %>% 
    dplyr::select(-Status) %>% 
    cor(method = "spearman")

}

# create df of potential regressors
df_hbox_reg <- df_hbox_impute %>% 
  dplyr::select(c(Status, `Hematocrit (%)`, `Lactate (mmol/L)`, `[HbO2]`, `[HbO2]α`, `[Hb]`, `[Hb]α`, `HbO2%`, `HbO2% α`, `[HbT]`, `[HbT]α`)) %>% 
  dplyr::filter(`[Hb]α` < 3)

m_cor_hv <- df_hbox_reg %>% f_calc_cor(status = "HV")

m_cor_um <- df_hbox_reg %>% f_calc_cor(status = "UM")

m_cor_cm <- df_hbox_reg %>% f_calc_cor(status = "CM")

# correlation matrix

# write plot function
f_plot_cor <- function(mat) {

  melt(mat) %>%

    mutate_if(is.factor, factor, levels = c("Lactate (mmol/L)", "Hematocrit (%)", "[HbO2]", "[Hb]", "HbO2%", "[HbT]", "[HbO2]α", "[Hb]α", "HbO2% α", "[HbT]α")) %>%

    ggplot(aes(x = X1, y = X2, fill = value, label = round(value, 2))) + 
    geom_tile() +
    scale_fill_gradient2(low = "#4575b4", high = "#d73027") +
    geom_text(size = 9) +

    xlab("") +
    ylab("") +
    theme_bw() +
    theme(text = element_text(family = "Arial"),
          axis.text.x = element_text(angle = 45, hjust = 0.9),
          axis.text = element_text(size = 35),
          legend.text = element_text(size = 40),
          legend.title = element_text(size = 50),
          legend.key.size = unit(2, 'cm'),
          plot.title = element_text(size = 30)
    )

}

  # extract legend
  p_leg <- m_cor_hv %>% f_plot_cor() %>% get_legend()

  # create plots
  m_cor_hv %>% f_plot_cor() + ggtitle("Healthy volunteers") + theme(legend.position = "none")

  m_cor_um %>% f_plot_cor() + ggtitle("Uncomplicated malaria patients") + theme(legend.position = "none")

  m_cor_cm %>% f_plot_cor() + ggtitle("Cerebral malaria patients") + theme(legend.position = "none")

  as_ggplot(p_leg)

```

**Figure S1.** Shown are the spearman correlations among variables measured by Oxiplex and their associated alpha values. Red indicates a positive correlation; blue indicates a negative correlation.

### Supplementary figure 2: Logistic regression curves of potential regressors
```{r fig_s2, fig.height = 12}

# pivot longer and change factor levels
df_hbox_reg_long <- df_hbox_reg %>% 
  pivot_longer(2:11, names_to = "variable", values_to = "value") %>% 
  mutate(variable = factor(variable, levels = c("Lactate (mmol/L)", "Hematocrit (%)", "[HbO2]", "[HbO2]α", "[Hb]", "[Hb]α", "HbO2%", "HbO2% α" ,"[HbT]", "[HbT]α")))

# plot as individual regression curves
df_hbox_reg_long %>% mutate(Status = as.numeric(ifelse(Status == "CM", 1, 0))) %>%

  ggplot(aes(x = value, y = Status)) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_wrap(~variable, scales = "free_x", nrow = 5, ncol = 2) +

  ylab("Status: 1 = CM, 0 = UM") +
  xlab("") +

  theme_bw() +
  theme(text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20),
        strip.text = element_text(size = 20))

```

**Figure S2.** Plotted are the individual logistic regression curves of individual variables tested for their discriminatory power between UM and CM. On the y-axis, 1 represents CM while 0 represents UM. A steeper curve indicates higher discriminatory power