### 2021-07-09 figures.Rmd
updated figures with new data

---
title: "Dysregulation of Cerebral Hemoglobin Content in Malawian Children with Malaria"
author: "Rachel Smith"
date: "`r Sys.Date()`"
output: html_document
---

<style type="text/css">
  body{
  font-family: Arial;
  font-size: 12pt;
}
</style>

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "", fig.height = 6, fig.width = 10)

library(tidyverse)
library(DescTools)
library(janitor)
library(FSA)
library(factoextra)
library(ggrepel)
library(ggpubr)
library(tidymodels)
library(vip)
library(scales)
library(reshape)
library(extrafont)
library(grDevices)

load("glm_model.Rdata")
load("results.Rdata")
load("dfa_example.Rdata")
load("dfa_results.Rdata")

font_import()

```

```{r data}

# read in data

  df_hbox <- read.csv("Data/master_datatable.csv") %>% 
    as_tibble() %>% 
    dplyr::select(-1) %>% 
    mutate(Status = factor(Status, levels = c("HV", "UM", "CM")))

# pivot long

  df_hbox_long <- df_hbox %>% pivot_longer(cols = 4:ncol(.), names_to = "variable", values_to = "value")

```

### Table 1: Summary of clinical data by patient group

```{r table1}

# number of subjects 
df_status <- df_hbox %>% 
  count(Status) %>% 
  pivot_wider(1:2, names_from = "Status", values_from = "n") %>% 
  mutate(variable = "Number of subjects") %>% 
  dplyr::select(variable, everything()) %>% 
  mutate_if(is.numeric, as.character)

# Sex
df_sex <- df_hbox %>% 
  group_by(Status) %>% 
  count(Sex) %>% 
  dplyr::filter(Sex != "") %>% 
  pivot_wider(names_from = "Sex", values_from = "n") %>% 
  mutate(perc_fem = round(Female/(Female + Male)*100, 2)) %>%

  pivot_wider(c(1,4), names_from = "Status", values_from = "perc_fem") %>% 
  mutate(variable = "Sex (% female)") %>% 
  dplyr::select(variable, everything()) %>% 
  mutate_if(is.numeric, as.character)

# other numeric variables

df_clinical_res <- df_hbox %>%

  # dplyr::filter(Subject.ID..NIAID. != "TM1018") %>%

  group_by(Status) %>% 
  summarise(`Age (years)` = median(Age, na.rm = TRUE), age_q1 = quantile(Age, 0.25, na.rm = TRUE), age_q3 = quantile(Age, 0.75, na.rm = TRUE),

            `Temperature (°C)` = median(Temperature, na.rm = TRUE), temp_q1 = quantile(Temperature, 0.25, na.rm = TRUE), temp_q3 = quantile(Temperature, 0.75, na.rm = TRUE),
            `Heart rate` = median(HR, na.rm = TRUE), hr_q1 = quantile(HR, 0.25, na.rm = TRUE), hr_q3 = quantile(HR, 0.75, na.rm = TRUE),
            `Systolic BP` = median(BP.Systolic, na.rm = TRUE), sys_q1 = quantile(BP.Systolic, 0.25, na.rm = TRUE), sys_q3 = quantile(BP.Systolic, 0.75, na.rm = TRUE),
            `Diastolic BP` = median(BP.Diastolic, na.rm = TRUE), dia_q1 = quantile(BP.Diastolic, 0.25, na.rm = TRUE), dia_q3 = quantile(BP.Diastolic, 0.75, na.rm = TRUE),
            `Respiration rate` = median(RR, na.rm = TRUE), rr_q1 = quantile(RR, 0.25, na.rm = TRUE), rr_q3 = quantile(RR, 0.75, na.rm = TRUE),
            `O~2~ saturation` = median(O2.Sat, na.rm = TRUE), o2_q1 = quantile(O2.Sat, 0.25, na.rm = TRUE), o2_q3 = quantile(O2.Sat, 0.75, na.rm = TRUE),

            `Hematocrit (%)` = median(Hematocrit, na.rm = TRUE), hem_q1 = quantile(Hematocrit, 0.25, na.rm = TRUE), hem_q3 = quantile(Hematocrit, 0.75, na.rm = TRUE),
            `Lactate (mmol/L)` = median(Lactate, na.rm = TRUE), lac_q1 = quantile(Lactate, 0.25, na.rm = TRUE), lac_q3 = quantile(Lactate, 0.75, na.rm = TRUE),
            `Arginine (μmol/L)` = median(Arginine.umol.L, na.rm = TRUE), arg_q1 = quantile(Arginine.umol.L, 0.25, na.rm = TRUE), arg_q3 = quantile(Arginine.umol.L, 0.75, na.rm = TRUE)) %>% 
  mutate_if(is.numeric, round, 1) %>%

  mutate(`Age (years)` = paste0(`Age (years)`, " (", age_q1, ", ", age_q3, ")"),

         `Temperature (°C)` = paste0(`Temperature (°C)`, " (", temp_q1, ", ", temp_q3, ")"),
         `Heart rate` = paste0(`Heart rate`, " (", hr_q1, ", ", hr_q3, ")"),
         `Systolic BP` = paste0(`Systolic BP`, " (", sys_q1, ", ", sys_q3, ")"),
         `Diastolic BP` = paste0(`Diastolic BP`, " (", dia_q1, ", ", dia_q3, ")"),
         `Respiration rate` = paste0(`Respiration rate`, " (", rr_q1, ", ", rr_q3, ")"),
         `O~2~ saturation` = paste0(`O~2~ saturation`, " (", o2_q1, ", ", o2_q3, ")"),

         `Hematocrit (%)` = paste0(`Hematocrit (%)`, " (", hem_q1, ", ", hem_q3, ")"),
         `Lactate (mmol/L)` = paste0(`Lactate (mmol/L)`, " (", lac_q1, ", ", lac_q3, ")"),
         `Arginine (μmol/L)` = paste0(`Arginine (μmol/L)`, " (", arg_q1, ", ", arg_q3, ")")) %>% 
  dplyr::select(-c(age_q1, age_q3, 
                   temp_q1, temp_q3, hr_q1, hr_q3, sys_q1, sys_q3, dia_q1, dia_q3, rr_q1, rr_q3, o2_q1, o2_q3, 
                   hem_q1, hem_q3, lac_q1, lac_q3, arg_q1, arg_q3)) %>%

  unique() %>% 
  t() %>% 
  row_to_names(row_number = 1) %>% 
  as_tibble(rownames = "variable") %>% 
  add_row(df_status, .before = 1) %>% 
  add_row(df_sex, .before = 2) %>%

  dplyr::rename(" " = "variable",
                "Healthy volunteers" = "HV",
                "Uncomplicated malaria patients" = "UM",
                "Cerebral malaria patients" = "CM")

  # remove lactate data for HV
  df_clinical_res[11,]$`Healthy volunteers` <- "--"

  # create kable df
  df_clinical_res %>% knitr::kable(align = rep('c', 4))

```
**Table 1.** A summary of subject information and hematological data. Group patient data presented as median (quartile 1, quartile 3). Note: Sex not recorded for 3 CM patients; Lactate data removed for healthy volunteers due to inconsistent data.

### Figure 1: Cerebral tissue hemoglobin concentration and oxygen saturation

```{r fig1}

# extract variables of interest
voi <- df_results %>% colnames() %>% .[4:9]

# select from the long df
df_hbox_voi_long <- df_hbox_long %>% dplyr::filter(variable %in% voi)

# create color vector
my_col <- c("CM" = "#00BFC4", "HV" = "#7CAE00", "UM" = "#F8766D")

# plot
df_hbox_voi_long %>% dplyr::filter(!(grepl("alpha", variable))) %>%

  ggplot(aes(x = Status, y = value)) +
  geom_violin(aes(color = Status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = Status), size = 0.2, width = 0.5) +
  facet_wrap(~variable, scales = "free_y") +

  scale_color_manual(values = my_col) +
  theme_bw() +
  theme(legend.position = "none")  +

  theme(text = element_text(family = "Arial"),
        strip.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        axis.title.y = element_blank(),
        axis.text = element_text(size = 15))

```

```{r fig1.pvals, eval = FALSE}

# hemoglobin concentration
kruskal.test(avg_Hb_conc ~ Status, data = df_hbox)
DescTools::DunnTest(avg_Hb_conc ~ Status, data = df_hbox, method = "BH")

# hemoglobin oxygen saturation
kruskal.test(avg_Hb_o2sat ~ Status, data = df_hbox)

```

**Figure 1.** Shown are the variables measured by Oxiplex, plotted by patient status. Cerebral tissue hemoglobin concentration levels were different between cerebral malaria patients and the other two groups, but did not differ between uncomplicated malaria patients and healthy volunteers (Kruskal-Wallis adjusted p-value = 5.0x10^-7^; Dunn Test post-hoc adjusted p-values: CM vs HV = 3.2x10^-6^, CM vs UM = 2.8x10^-4^, HV vs UM = 0.87). Levels of cerebral hemoglobin O~2~ saturation did not differ among groups (Kruskal-Wallis adjusted p-value = 0.47).

### Figure 2: Detrended fluctuation analysis results

```{r fig2}

df_hbox_voi_long %>% dplyr::filter(grepl("second_alpha", variable)) %>%

  ggplot(aes(x = Status, y = value)) +
  geom_violin(aes(color = Status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = Status), size = 0.2, width = 0.5) +

  geom_hline(yintercept = 0.45, alpha = 0.4, lty = 2) +
  geom_hline(yintercept = 0.55, alpha = 0.4, lty = 2) +
  geom_hline(yintercept = 1.0, alpha = 0.4, lty = 2) +
  geom_hline(yintercept = 1.5, alpha = 0.4, lty = 2) +

  facet_wrap(~variable, scales = "free_y") +

  ylim(0, 1.5) +
  scale_color_manual(values = my_col) +
  theme_bw() +

  theme(text = element_text(family = "Arial"),
        strip.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        legend.position = "none")

```

```{r fig2.pvals, eval = FALSE}

# hemoglobin concentration alpha
kruskal.test(Hb_conc_second_alpha ~ Status, data = df_hbox)
DescTools::DunnTest(Hb_conc_second_alpha ~ Status, data = df_hbox, method = "BH")

# hemoglobin oxygen saturation alpha
kruskal.test(Hb_o2sat_second_alpha ~ Status, data = df_hbox)
DescTools::DunnTest(Hb_o2sat_second_alpha ~ Status, data = df_hbox, method = "BH")

```

**Figure 2.** Shown are the alpha values associated with the hemoglobin concentration and oxygen saturation signals measured by Oxiplex. Differences among cerebral tissue hemoglobin concentration alpha values were significant by Kruskal-Wallis (adjusted p-value = 7.16x10^-4^), and differences between uncomplicated malaria patients and the other two patient groups were significant by post-hoc Dunn Test (adjusted p-values: CM vs HV = 0.10, CM vs UM = 4.7x10^-4^, HV vs UM = 0.03). Cerebral hemoglobin O~2~ saturation alpha values differed between cerebral malaria patients and the other two patient groups (Kruskal-Wallis adjusted p-value = 4.2x10^-5^; Dunn Test post-hoc adjusted p-values: CM vs HV = 7.5x10^-4^, CM vs UM = 3.5x10^-4^, HV vs UM = 0.32).

### Figure 3: Logistic regression model on CM vs UM

#### A. PCA
```{r fig3a, fig.height = 8}

# include variables of interest, remove healthy volunteers from analysis

df_hbox_voi <- df_hbox %>% dplyr::select(c("subject_id", "Status", "Hematocrit", "Lactate", "avg_Hb_conc", "Hb_conc_second_alpha", "avg_Hb_o2sat", "Hb_o2sat_second_alpha"))

# impute missing data with median by group
  f_impute <- function(x, na.rm = TRUE) (replace(x, is.na(x), median(x, na.rm = na.rm)))

  df_hbox_impute <- df_hbox_voi %>% 
    group_by(Status) %>% 
    mutate_at(3:ncol(.), f_impute) %>% 
    ungroup()

# remove healthy volunteers for logistic regression analysis
df_hbox_log_reg <- df_hbox_impute %>% dplyr::filter(Status != "HV")

# convert tibble to df
df_hbox_voi_df <- df_hbox_log_reg %>% 
  dplyr::select(-c(subject_id, Status)) %>% 
  as.data.frame()

# add rownames
rownames(df_hbox_voi_df) <- df_hbox_log_reg %>% .$subject_id

# compute PCA
my_pca <- prcomp(df_hbox_voi_df, scale = TRUE)

# define groups
groups <- as.factor(df_hbox_log_reg$Status)

# Eigenvalues
pca_eig_val <- get_eigenvalue(my_pca)

# two data frames for plotting
df_vars <- my_pca$rotation %>% as_tibble() %>% mutate(variable = rownames(my_pca$rotation))
df_pnts <- my_pca$x %>% as_tibble() %>% mutate(group = groups)

# plot

  # plot patients
  ggplot() +
  geom_point(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = group), size = 3) +
  stat_conf_ellipse(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = group, fill = group),
                    level = 0.95, npoint = 100, bary = TRUE, alpha = 0.1, geom = "polygon") +
  stat_mean(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = group, shape = group),
            na.rm = FALSE) +

  # plot variables
  geom_point(data = df_vars, aes(x = PC1*5, y = PC2*5)) +
  geom_text_repel(data = df_vars, aes(x = PC1*5, y = PC2*5, label = variable, size = 10), show.legend = FALSE) +
  geom_segment(data = df_vars, aes(x = 0, y = 0, xend = PC1*5, yend = PC2*5), arrow = arrow()) +

  # draw lines
  geom_vline(xintercept = 0, lty = 2, color = "black") +
  geom_hline(yintercept = 0, lty = 2, color = "black") +

  # design
  xlab(paste0("PC1: ", round(pca_eig_val$variance.percent[1], 2), "% of variance")) +
  ylab(paste0("PC2: ", round(pca_eig_val$variance.percent[2], 2), "% of variance")) +

  theme_bw() +
  theme(text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))

```

#### B. Confusion matrix

```{r fig3b_prep, eval = FALSE}

# select [HbT], [HbT] alpha, and hematocrit
  df_hbox_model <- df_hbox_log_reg %>% 
    dplyr::select(c(Status, Hematocrit, avg_Hb_conc, Hb_conc_second_alpha)) %>% 
    mutate(`1/10_Hb_conc_alpha` = Hb_conc_second_alpha*0.1, avg_Hb_conc_nm = avg_Hb_conc/1000) %>% 
    dplyr::select(-c(Hb_conc_second_alpha, avg_Hb_conc)) %>% 
    mutate(Status = as.factor(ifelse(Status == "CM", 1, 0)))

# resample
set.seed(234)
hbox_boot <- bootstraps(df_hbox_model, strata = Status, times = 1000)

# recipe
hbox_rec <- recipe(Status ~ ., data = df_hbox_model) %>% 
  step_BoxCox(all_predictors()) %>% 
  step_nzv(all_predictors())

# create workflow
hbox_wf <- workflow() %>% add_recipe(hbox_rec)

# logistic regression model specifications
glm_spec <- logistic_reg() %>% set_engine("glm")

# create model on 1000 different resamples
glm_res <- hbox_wf %>% 
  add_model(glm_spec) %>%
  fit_resamples(
    resamples = hbox_boot,
    metrics = metric_set(roc_auc, accuracy, sensitivity, specificity),
    control = tune::control_resamples(save_pred = TRUE, verbose = TRUE)
  )

# select best model to be used for evaluation
glm_best <- glm_res %>% select_best("roc_auc")

# finalize model
hbox_final <- hbox_wf %>% 
  add_model(glm_spec) %>% 
  finalize_workflow(glm_best) %>%
  fit(df_hbox_model) %>%
  pull_workflow_fit()

# save objects so model doesn't rerun every time we knit
save(glm_res, hbox_final, file = "glm_model.Rdata")

```

```{r fig3b_plot, fig.height = 8}

final_res_conf_mat <- glm_res %>% 
  conf_mat_resampled() %>% 
  mutate(Prediction = ifelse(Prediction == 1, "CM", "UM")) %>% 
  mutate(Truth = ifelse(Truth == 1, "CM", "UM"))

final_res_conf_mat %>% 
  group_by(Truth) %>% 
  summarise(total_true = sum(Freq)) %>% 
  right_join(final_res_conf_mat) %>% 
  mutate(`Percent occurence` = round(Freq/total_true*100, 2)) %>%

  # plot
  ggplot(aes(x = Truth, y = Prediction, fill = `Percent occurence`, label = `Percent occurence`)) +
  geom_tile() +
  geom_text(size = 6) +
  scale_fill_gradient2(low = "#4575b4", high = "#d73027") +

  theme_bw() +
  theme(text = element_text(family = "Arial"),
        axis.title = element_text(size = 25),
        axis.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 15))

```

**Figure 3.** Logistic regression model discriminating CM vs UM patients using cerebral tissue hemoglobin concentration, cerebral tissue hemoglobin concentration alpha, and hematocrit as the regressors **(A)** PCA of patients with CM and UM as determined by NIRS variables. Color indicates patient status, ellipses show median and 95% confidence interval on the cluster. **(B)** Confusion matrix representing the results of the logistic regression model averaged across 1000 bootstrap resamples. Percent occurrence indicates percentage times that the event (shown by the x- and y-axes) occurred.

### Table 2. Logistic regression model evaluation

#### A. Model metrics
```{r table2a}

collect_metrics(glm_res) %>% 
  dplyr::select(.metric, mean, std_err) %>% 
  mutate(mean = mean*100, std_err = std_err*100) %>% 
  mutate(Metric = c("Accuracy", "ROC AUC", "Sensitivity", "Specificity")) %>% 
  dplyr::select(c("Metric", "mean", "std_err")) %>%

  dplyr::rename("Mean" = "mean", "Standard error" = "std_err") %>% 
  mutate_if(is.numeric, round, 2) %>% 
  as.data.frame() %>%

  knitr::kable(align = rep('c', 4))

```

#### B. Coefficients
```{r table2b}

hbox_final %>% 
  tidy(exponentiate = TRUE) %>% 
  mutate(estimate = ifelse(estimate < 10^-2, scientific(estimate, digits = 3), round(estimate, 2))) %>% 
  mutate(p.value = scientific(p.value, digits = 3),
         std.error = round(std.error, 2),
         statistic = round(statistic, 2)) %>% 
  dplyr::rename("Term" = "term", "Odds ratio" = "estimate", "Standard error" = "std.error", 
                "Statistic" = "statistic", "p-value" = "p.value") %>%

  as.data.frame() %>%
  knitr::kable(align = rep('c', 5))

```

**Table 2.** Evaluation of the best logistic regression model select from 1000.

### Figure 4: Relationships among hemoglobin concentration and hematocrit
```{r fig4, fig.height = 8}

df_hbox %>% filter(Status != "HV") %>%

  ggplot(aes(x = avg_Hb_conc, y = Hematocrit)) +
  geom_point(aes(color = Status), alpha = 0.7, size = 3) +

  xlab("Hemoglobin concentration (μM)") +
  ylab("Hematocrit (%)") +

  theme_bw() +
  theme(text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))

```

**Figure 4.**

### Supplementary figure 1: Spearman correlations of potential regressors
```{r fig_s1, fig.show = "hold", out.width = "50%", fig.height = 12, fig.width = 15}

# calculate spearman correlations in each group

f_calc_cor <- function(df, status) {

  df %>%
    dplyr::filter(Status == status) %>% 
    dplyr::select(-c(subject_id, Status)) %>% 
    cor(method = "spearman")

}

# create run correlations of potential regressors
m_cor_hv <- df_hbox_impute %>% f_calc_cor(status = "HV")

m_cor_um <- df_hbox_impute %>% f_calc_cor(status = "UM")

m_cor_cm <- df_hbox_impute %>% f_calc_cor(status = "CM")

# correlation matrix

# write plot function
f_plot_cor <- function(mat) {

  melt(mat) %>%

    mutate_if(is.factor, factor, levels = c("Lactate", "Hematocrit", "avg_Hb_o2sat", "avg_Hb_conc", "Hb_o2sat_second_alpha", "Hb_conc_second_alpha")) %>%

    ggplot(aes(x = X1, y = X2, fill = value, label = round(value, 2))) + 
    geom_tile() +
    scale_fill_gradient2(low = "#4575b4", high = "#d73027") +
    geom_text(size = 9) +

    xlab("") +
    ylab("") +
    theme_bw() +
    theme(text = element_text(family = "Arial"),
          axis.text.x = element_text(angle = 45, hjust = 0.9),
          axis.text = element_text(size = 35),
          legend.text = element_text(size = 40),
          legend.title = element_text(size = 50),
          legend.key.size = unit(2, 'cm'),
          plot.title = element_text(size = 30)
    )

}

  # extract legend
  p_leg <- m_cor_hv %>% f_plot_cor() %>% get_legend()

  # create plots
  m_cor_hv %>% f_plot_cor() + ggtitle("Healthy volunteers") + theme(legend.position = "none")

  m_cor_um %>% f_plot_cor() + ggtitle("Uncomplicated malaria patients") + theme(legend.position = "none")

  m_cor_cm %>% f_plot_cor() + ggtitle("Cerebral malaria patients") + theme(legend.position = "none")

  as_ggplot(p_leg)

```

**Figure S1.** Shown are the spearman correlations among hematocrit, lactate, hemoglobin concentration and its alpha, and hemoglobin oxygen saturation and its alpha. Red indicates a positive correlation; blue indicates a negative correlation.

### Supplementary figure 2: Logistic regression curves of potential regressors
```{r fig_s2, fig.height = 12}

# pivot longer and change factor levels
df_hbox_reg_long <- df_hbox_impute %>% 
  mutate(Status = as.numeric(ifelse(Status == "CM", 1, 0))) %>% 
  pivot_longer(3:ncol(.), names_to = "variable", values_to = "value") %>% 
  mutate(variable = factor(variable, levels = c("Lactate", "Hematocrit", "avg_Hb_o2sat", "avg_Hb_conc", "Hb_o2sat_second_alpha", "Hb_conc_second_alpha")))

# plot as individual regression curves
df_hbox_reg_long %>%

  ggplot(aes(x = value, y = Status)) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_wrap(~variable, scales = "free_x", nrow = 5, ncol = 2) +

  ylab("Status: 1 = CM, 0 = UM") +
  xlab("") +

  theme_bw() +
  theme(text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20),
        strip.text = element_text(size = 20))

```

**Figure S2.** Plotted are the individual logistic regression curves of individual variables tested for their discriminatory power between UM and CM. On the y-axis, 1 represents CM while 0 represents UM. A steeper curve indicates higher discriminatory power.

### Supplementary figure 3: Detrended fluctuation analysis

#### A. Pre-processed signal
```{r fig_s3a}

df_thc_cumsum1 %>%

  ggplot(aes(x = Time, y = sg_filt)) +
  geom_line() +

  labs(y = "Hemoglobin concentration (μM)", x = "Time (s)") +
  ggtitle("Pre-processed signal") +
  theme_bw() +
  theme(text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))

```

#### B. Cumulative sum
```{r fig_s3b}

df_thc_cumsum1 %>%

  ggplot(aes(x = Time, y = cumsum)) +
  geom_line() +
  geom_vline(xintercept = 1201, lty = 2, color = "black") +
  geom_vline(xintercept = 1210, lty = 2, color = "black") +

  labs(y = "Cumulative sum", x = "Time (s)") +
  ggtitle("Cumulative signal") +
  theme_bw() +
  theme(text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))

```

#### C. Window detrend
```{r fig_s3c, fig.show = "hold", out.width = "50%"}

fit <- lm(cumsum ~ Time, data = df_thc_cumsum1 %>% dplyr::filter(Time > 1201 & Time < 1210))

# Window 
df_thc_cumsum1 %>% dplyr::filter(Time > 1201 & Time < 1210) %>%

  ggplot(aes(x = Time, y = cumsum)) +
  geom_line(size = 2) +
  geom_abline(intercept = coef(fit)[1], slope = coef(fit)[2], color = "black", lty = 2, size = 2) +

  labs(y = "Cumulative sum", x = "Time (s)") +
  ggtitle("Cumulative signal") +
  theme_bw() +
  theme(text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20),
        title = element_text(size = 25))

# Detrended
df_thc_cumsum1 %>% dplyr::filter(Time > 1201 & Time < 1210) %>%

  ggplot(aes(x = Time, y = fit$residuals)) +
  geom_line(size = 2) +
  geom_hline(yintercept = 0, color = "black", lty = 2, size = 2) +

  labs(y = "", x = "Time (s)") +
  ggtitle("Detrended cumulative signal") +
  theme_bw() +
  theme(text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20),
        title = element_text(size = 25))

```

#### D. Results plot
```{r fig_s3d}

# find slope of segmented line
library(segmented)
df1 <- l_dfa_thc[[1]][[1]] %>% mutate(log10_avg_fluct = log10(avg_fluctuation), log10_window_size = log10(window_size))
fit1 <- lm(log10_avg_fluct ~ log10_window_size, data = df1) 
segfit1 <- segmented(fit1)
slopes1 <- coef(segfit1) # the coefficients are the differences in slope in comparison to the previous slope

# first line: 
#y = b0 + b1*x; y = intercept1 + slope1 * x

# second line:
# y = c0 + c1*x; y = intercept2 + slope2 * x

# At the breakpoint (break1), the segments b and c intersect: b0 + b1*x = c0 + c1*x
b0 <- slopes1[[1]]
b1 <- slopes1[[2]]

c1 <- slopes1[[2]] + slopes1[[3]]
break1 <- segfit1$psi[[2]]

# Solve for c0 (intercept of second segment):
c0 <- b0 + b1 * break1 - c1 * break1

df1 %>%

  ggplot(aes(x = log10_window_size, y = log10_avg_fluct)) +
  geom_point(size = 3, shape = 1) +
  geom_abline(intercept = c0, slope = c1, color = "black", lty = 2) +
  annotate("text", x = -0.5, y = 2, label = paste0("alpha = ", round(c1, 2)), size = 8) +

  labs(x = "log10(window length)", y = "log10(average fluctuation)") +
  ggtitle("DFA results") +

  theme_bw() +
  theme(text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))

```