### 2021-07-25 cm_follow_up_figures.Rmd
change order of variables in figures; export table to csv

---
title: "Cerebral malaria follow-up visit data"
author: "Rachel Smith"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "", fig.height = 6, fig.width = 10)

library(tidyverse)
library(extrafont)

load("results.Rdata")
load("follow_up_results.Rdata")

font_import()

```

```{r data}

# combine visit 1 with follow-ups
df_results_all <- df_results %>% 
  mutate(subject_id = substr(subject_id, start = 1, stop = 6)) %>% 
  left_join(df_fu_results %>% select(-c(number, Status)), by = "subject_id")

# reorganize for plotting
df_results_all_plot <- df_results_all %>% 
  dplyr::filter(subject_id %in% df_fu_results$subject_id) %>% 
  pivot_longer(cols = 4:ncol(.), names_to = "variable", values_to = "value") %>% 
  mutate(follow_up = ifelse(grepl("fu", variable), "follow-up", "initial visit")) %>% 
  mutate(follow_up = factor(follow_up, levels = c("initial visit", "follow-up")))

# descriptive statistics

# THC
df_thc_compare <- df_results_all_plot %>% 
  dplyr::filter(variable %in% c("Hb_conc_overall_alpha", "Hb_conc_fu_alpha"))

thc_fu_pval <- wilcox.test(value ~ follow_up, data = df_thc_compare) %>% .$p.value

# O2 sat
df_o2sat_compare <- df_results_all_plot %>% 
  dplyr::filter(variable %in% c("Hb_o2sat_overall_alpha", "Hb_o2sat_fu_alpha"))

o2sat_fu_pval <- wilcox.test(value ~ follow_up, data = df_o2sat_compare) %>% .$p.value

```

### Figure 5: Cerebral malaria patient follow-up visits

```{r fig5}

# plot O2 sat
df_results_all_plot %>% 
  dplyr::filter(!grepl("avg|second", variable)) %>% 
  mutate(measurement = ifelse(grepl("conc", variable), "THC", "O2_sat")) %>% 
  mutate(measurement = factor(measurement, levels = c("O2_sat", "THC"))) %>% 
  mutate(pvals = ifelse(measurement == "THC", round(thc_fu_pval, 3), round(o2sat_fu_pval, 3))) %>%

  ggplot(aes(x = follow_up, y = value)) +
  geom_violin(aes(color = follow_up)) +
  geom_point(shape = 1, size = 2) +
  stat_summary(fun = "median", geom = "crossbar", aes(x = follow_up, color = follow_up), size = 0.2, width = 0.5) +
  geom_path(aes(group = subject_id), lty = 2, alpha = 0.7) +

  geom_segment(aes(x = 1, y = 1.4, xend = 2, yend = 1.4), size = 0.3) +
  geom_segment(aes(x = 1, y = 1.4, xend = 1, yend = 1.35), size = 0.3) +
  geom_segment(aes(x = 2, y = 1.4, xend = 2, yend = 1.35), size = 0.3) +
  geom_text(x = 1.5, y = 1.43, aes(label = paste0("p = ", pvals))) +

  facet_wrap(~measurement) +

  ylim(0.5, 1.5) +
  labs(y = "alpha", x = "") +

  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "Arial"),
        strip.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        axis.text = element_text(size = 15),
        legend.position = "none")

```