### 2021-11-05 final_figures.Rmd
scripts, no edits

---
title: "Cerebral oxygenation manuscript figures"
date: "`r Sys.Date()`"
output: html_document
---

<style type="text/css">
  body{
  font-family: Arial;
  font-size: 12pt;
}
</style>

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "", fig.height = 6, fig.width = 10)

# Load libraries
library(tidyverse)
library(DescTools)
library(janitor)
library(FSA)
library(factoextra)
library(ggrepel)
library(ggpubr)
library(tidymodels)
library(vip)
library(scales)
library(reshape)
library(extrafont)
library(grDevices)

# Load data
load("psd_plot_data.Rdata")
load("glm_model.Rdata")
load("results.Rdata")
load("roc_curves.Rdata")
load("dfa_example.Rdata")
load("dfa_results.Rdata")
load("all_dfa_results.Rdata")

font_import()

```

```{r data}

# read in data

  df_hbox <- read.csv("Data/master_datatable.csv") %>% 
    as_tibble() %>% 
    dplyr::select(-1) %>% 
    mutate(Status = replace(Status, Status == "HV", "HC")) %>% 
    mutate(Status = factor(Status, levels = c("HC", "UM", "CM")))

# pivot long

  df_hbox_long <- df_hbox %>% pivot_longer(cols = 7:ncol(.), names_to = "variable", values_to = "value")

```

### Table 1. Baseline clinical characteristics

```{r table1}

# number of subjects 
df_status <- df_hbox %>% 
  count(Status) %>% 
  pivot_wider(1:2, names_from = "Status", values_from = "n") %>% 
  mutate(variable = "Number of subjects") %>% 
  dplyr::select(variable, everything()) %>% 
  mutate_if(is.numeric, as.character)

# Sex
df_sex <- df_hbox %>% 
  group_by(Status) %>% 
  count(Sex) %>% 
  dplyr::filter(Sex != "") %>% 
  pivot_wider(names_from = "Sex", values_from = "n") %>% 
  mutate(perc_fem = round(Female/(Female + Male)*100, 2)) %>%

  pivot_wider(c(1,4), names_from = "Status", values_from = "perc_fem") %>% 
  mutate(variable = "Sex (% female)") %>% 
  dplyr::select(variable, everything()) %>% 
  mutate_if(is.numeric, as.character)

# other numeric variables

df_clinical_res <- df_hbox %>%

  # dplyr::filter(Subject.ID..NIAID. != "TM1018") %>%

  group_by(Status) %>% 
  summarise(`Age (years)` = median(Age, na.rm = TRUE), age_q1 = quantile(Age, 0.25, na.rm = TRUE), age_q3 = quantile(Age, 0.75, na.rm = TRUE),

            `Temperature (°C)` = median(Temperature, na.rm = TRUE), temp_q1 = quantile(Temperature, 0.25, na.rm = TRUE), temp_q3 = quantile(Temperature, 0.75, na.rm = TRUE),
            `Heart rate` = median(HR, na.rm = TRUE), hr_q1 = quantile(HR, 0.25, na.rm = TRUE), hr_q3 = quantile(HR, 0.75, na.rm = TRUE),
            `Systolic BP` = median(BP.Systolic, na.rm = TRUE), sys_q1 = quantile(BP.Systolic, 0.25, na.rm = TRUE), sys_q3 = quantile(BP.Systolic, 0.75, na.rm = TRUE),
            `Diastolic BP` = median(BP.Diastolic, na.rm = TRUE), dia_q1 = quantile(BP.Diastolic, 0.25, na.rm = TRUE), dia_q3 = quantile(BP.Diastolic, 0.75, na.rm = TRUE),
            `Respiration rate` = median(RR, na.rm = TRUE), rr_q1 = quantile(RR, 0.25, na.rm = TRUE), rr_q3 = quantile(RR, 0.75, na.rm = TRUE),
            `O~2~ saturation` = median(O2.Sat, na.rm = TRUE), o2_q1 = quantile(O2.Sat, 0.25, na.rm = TRUE), o2_q3 = quantile(O2.Sat, 0.75, na.rm = TRUE),

            `Hematocrit (%)` = median(Hematocrit, na.rm = TRUE), hem_q1 = quantile(Hematocrit, 0.25, na.rm = TRUE), hem_q3 = quantile(Hematocrit, 0.75, na.rm = TRUE),
            `Lactate (mmol/L)` = median(Lactate, na.rm = TRUE), lac_q1 = quantile(Lactate, 0.25, na.rm = TRUE), lac_q3 = quantile(Lactate, 0.75, na.rm = TRUE),
            `Arginine (μmol/L)` = median(Arginine.umol.L, na.rm = TRUE), arg_q1 = quantile(Arginine.umol.L, 0.25, na.rm = TRUE), arg_q3 = quantile(Arginine.umol.L, 0.75, na.rm = TRUE)) %>% 
  mutate_if(is.numeric, round, 1) %>%

  mutate(`Age (years)` = paste0(`Age (years)`, " (", age_q1, ", ", age_q3, ")"),

         `Temperature (°C)` = paste0(`Temperature (°C)`, " (", temp_q1, ", ", temp_q3, ")"),
         `Heart rate` = paste0(`Heart rate`, " (", hr_q1, ", ", hr_q3, ")"),
         `Systolic BP` = paste0(`Systolic BP`, " (", sys_q1, ", ", sys_q3, ")"),
         `Diastolic BP` = paste0(`Diastolic BP`, " (", dia_q1, ", ", dia_q3, ")"),
         `Respiration rate` = paste0(`Respiration rate`, " (", rr_q1, ", ", rr_q3, ")"),
         `O~2~ saturation` = paste0(`O~2~ saturation`, " (", o2_q1, ", ", o2_q3, ")"),

         `Hematocrit (%)` = paste0(`Hematocrit (%)`, " (", hem_q1, ", ", hem_q3, ")"),
         `Lactate (mmol/L)` = paste0(`Lactate (mmol/L)`, " (", lac_q1, ", ", lac_q3, ")"),
         `Arginine (μmol/L)` = paste0(`Arginine (μmol/L)`, " (", arg_q1, ", ", arg_q3, ")")) %>% 
  dplyr::select(-c(age_q1, age_q3, 
                   temp_q1, temp_q3, hr_q1, hr_q3, sys_q1, sys_q3, dia_q1, dia_q3, rr_q1, rr_q3, o2_q1, o2_q3, 
                   hem_q1, hem_q3, lac_q1, lac_q3, arg_q1, arg_q3)) %>%

  unique() %>% 
  t() %>% 
  row_to_names(row_number = 1) %>% 
  as_tibble(rownames = "variable") %>% 
  add_row(df_status, .before = 1) %>% 
  add_row(df_sex, .before = 2) %>%

  dplyr::rename(" " = "variable",
                "Healthy children" = "HC",
                "Uncomplicated malaria" = "UM",
                "Cerebral malaria" = "CM")

  # remove lactate data for HC
  df_clinical_res[11,]$`Healthy children` <- "--"

  # export to CSV
  # write.csv(df_clinical_res, file = "Outputs/table_1.csv")

  # create kable df
  df_clinical_res %>% knitr::kable(align = rep('c', 4))

```
**Table 1.** A summary of subject information and hematological data. Group patient data presented as median (quartile 1, quartile 3). Note: Sex not recorded for 3 CM patients; Lactate data removed for healthy volunteers due to inconsistent data.

### Fig 0. (optional) Power spectral density curves

```{r fig0}

# plot
f_plot_psd <- function(df, status, upper_lim) {

  df %>% 
    dplyr::filter(Status == status) %>% 
    dplyr::filter(freq > 0.01 & freq < upper_lim) %>% 
    ggplot(aes(x = freq, y = med_power)) +
    geom_line() +
    geom_vline(xintercept = 0.1, color = "red", lty = 2) +
    facet_grid(tissue ~ variable, scales = "free_y") +
    labs(x = "Frequency", y = "Median power across patients") +
    theme_bw()

}

## 0-1 Hz

  # HC
  f_plot_psd(df_psd, "HC", 1) + ggtitle("Healthy children")

  # UM
  f_plot_psd(df_psd, "UM", 1) + ggtitle("Uncomplicated malaria")

  # CM
  f_plot_psd(df_psd, "CM", 1) + ggtitle("Cerebral malaria")

## 0-3 Hz

  # HC
  f_plot_psd(df_psd, "HC", 3) + ggtitle("Healthy children")

  # UM
  f_plot_psd(df_psd, "UM", 3) + ggtitle("Uncomplicated malaria")

  # CM
  f_plot_psd(df_psd, "CM", 3) + ggtitle("Cerebral malaria")

```

### Fig 1. Hb_oxy and Hb_tot in the brain and muscle

```{r fig1}

# extract variables of interest
voi <- df_all_dfa_results %>% colnames() %>% .[-(1:2)]

# select from the long df
df_hbox_voi_long <- df_hbox_long %>% 
  dplyr::filter(variable %in% voi & !grepl("fu", variable)) %>%

  mutate(tissue = as.factor(ifelse(grepl("muscle", variable), "muscle", "brain")),
         measurement = factor(ifelse(grepl("conc", variable), "Hb_tot", "Hb_oxy"), levels = c("Hb_oxy", "Hb_tot")),
         metric = case_when(

           grepl("overall_alpha|conc_alpha|o2sat_alpha", variable) ~ "alpha",
           grepl("second_alpha", variable) ~ "second_alpha",
           TRUE ~ "mean"

         )) %>% 
  dplyr::select(-variable)

# create color vector
my_col <- c("CM" = "#00BFC4", "HC" = "#7CAE00", "UM" = "#F8766D")

# plot
df_hbox_voi_long %>% dplyr::filter(!(grepl("alpha", metric))) %>%

  ggplot(aes(x = Status, y = value)) +
  geom_violin(aes(color = Status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1, size = 2) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = Status), size = 0.2, width = 0.5) +
  facet_grid(measurement ~ tissue, scales = "free_y") +

  scale_color_manual(values = my_col) +
  labs(x = "") +
  theme_bw() +
  theme(legend.position = "none")  +

  theme(# text = element_text(family = "Arial"),
        strip.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        axis.title.y = element_blank(),
        axis.text = element_text(size = 15))

```

### Fig 2. Hb_oxy alpha and Hb_tot alpha in the brain and muscle

```{r fig2}

# plot
df_hbox_voi_long %>% dplyr::filter((grepl("second_alpha", metric))) %>%

  ggplot(aes(x = Status, y = value)) +
  geom_violin(aes(color = Status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1, size = 2) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = Status), size = 0.2, width = 0.5) +
  facet_grid(measurement ~ tissue) +

  geom_hline(yintercept = 0.45, alpha = 0.4, lty = 2) +
  geom_hline(yintercept = 0.55, alpha = 0.4, lty = 2) +
  geom_hline(yintercept = 1.0, alpha = 0.4, lty = 2) +
  geom_hline(yintercept = 1.5, alpha = 0.4, lty = 2) +

  ylim(0, 1.5) +
  scale_color_manual(values = my_col) +
  labs(x = "") +
  theme_bw() +
  theme(legend.position = "none")  +

  theme(# text = element_text(family = "Arial"),
        strip.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        axis.title.y = element_blank(),
        axis.text = element_text(size = 15))

```

### Fig 3. Logistic regression modeling of uncomplicated and cerebral malaria subjects

```{r fig3_data}

# include variables of interest

df_hbox_log_reg <- df_hbox %>% 
  dplyr::select(c("subject_id", "Status", "Hematocrit", "Lactate", "avg_Hb_conc", "Hb_conc_second_alpha", "avg_Hb_o2sat", "Hb_o2sat_second_alpha"))

# impute missing data with median by group
  f_impute <- function(x, na.rm = TRUE) (replace(x, is.na(x), median(x, na.rm = na.rm)))

  df_hbox_impute <- df_hbox_log_reg %>% 
    group_by(Status) %>% 
    mutate_at(3:ncol(.), f_impute) %>% 
    ungroup()

```

#### A. Correlations of variables within subject groups

```{r fig3a, fig.show = "hold", out.width = "50%", fig.height = 12, fig.width = 15}

# calculate spearman correlations in each group

f_calc_cor <- function(df, status) {

  m <- df %>%
    dplyr::filter(Status == status) %>% 
    dplyr::select(-c(subject_id, Status)) %>% 
    cor(method = "spearman")

  diag(m) <- 0

  m
}

# create run correlations of potential regressors
m_cor_um <- df_hbox_impute %>% f_calc_cor(status = "UM")

m_cor_cm <- df_hbox_impute %>% f_calc_cor(status = "CM")

# correlation matrix

# write plot function
f_plot_cor <- function(mat) {

  melt(mat) %>%

    mutate_if(is.factor, factor, levels = c("Lactate", "Hematocrit", "avg_Hb_o2sat", "avg_Hb_conc", "Hb_o2sat_second_alpha", "Hb_conc_second_alpha")) %>%

    ggplot(aes(x = X1, y = X2, fill = value, label = ifelse(value != 0, round(value, 2), NA))) + 
    geom_tile() +
    scale_fill_gradient2(low = "#4575b4", high = "#d73027") +
    geom_text(size = 9) +

    xlab("") +
    ylab("") +
    theme_bw() +
    theme(# text = element_text(family = "Arial"),
          axis.text.x = element_text(angle = 45, hjust = 1),
          axis.text = element_text(size = 35),
          legend.text = element_text(size = 40),
          legend.title = element_text(size = 50),
          legend.key.size = unit(2, 'cm'),
          plot.title = element_text(size = 30)
    )

}

  # extract legend
  p_leg <- m_cor_um %>% f_plot_cor() %>% get_legend()

  # create plots
  m_cor_um %>% f_plot_cor() + ggtitle("Uncomplicated malaria") + theme(legend.position = "none")

  m_cor_cm %>% f_plot_cor() + ggtitle("Cerebral malaria") + theme(legend.position = "none")

  as_ggplot(p_leg)

```

#### B. Principal component analysis

```{r fig3b}
# remove healthy volunteers for logistic regression analysis
df_hbox_impute <- df_hbox_impute %>% dplyr::filter(Status != "HC")

# convert tibble to df
df_hbox_df <- df_hbox_impute %>% 
  dplyr::select(-c(subject_id, Status)) %>% 
  as.data.frame()

# add rownames
rownames(df_hbox_df) <- df_hbox_impute %>% .$subject_id

# compute PCA
my_pca <- prcomp(df_hbox_df, scale = TRUE)

# define groups
groups <- as.factor(df_hbox_impute$Status)

# Eigenvalues
pca_eig_val <- get_eigenvalue(my_pca)

# two data frames for plotting
df_vars <- my_pca$rotation %>% as_tibble() %>% mutate(variable = rownames(my_pca$rotation))
df_pnts <- my_pca$x %>% as_tibble() %>% mutate(group = groups)

# plot

  # plot patients
  ggplot() +
  geom_point(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = group), size = 3) +
  stat_conf_ellipse(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = group, fill = group),
                    level = 0.95, npoint = 100, bary = TRUE, alpha = 0.1, geom = "polygon") +
  stat_mean(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = group, shape = group),
            na.rm = FALSE) +

  # plot variables
  # geom_point(data = df_vars, aes(x = PC1*5, y = PC2*5)) +
  # geom_text_repel(data = df_vars, aes(x = PC1*5, y = PC2*5, label = variable, size = 10), show.legend = FALSE) +
  geom_segment(data = df_vars, aes(x = 0, y = 0, xend = PC1*5, yend = PC2*5), arrow = arrow()) +

  # draw lines
  geom_vline(xintercept = 0, lty = 2, color = "black") +
  geom_hline(yintercept = 0, lty = 2, color = "black") +

  # design
  xlab(paste0("PC1: ", round(pca_eig_val$variance.percent[1], 2), "% of variance")) +
  ylab(paste0("PC2: ", round(pca_eig_val$variance.percent[2], 2), "% of variance")) +

  theme_bw() +
  theme(# text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))

```

#### C. Coefficients
```{r table2c}

df_CI <- confint(hbox_final$fit) %>% exp() %>% as_tibble(rownames = "term")

table2c <- hbox_final %>%
  tidy(exponentiate = TRUE) %>%
  left_join(df_CI, by = "term") %>%
  mutate(estimate = ifelse(estimate < 10^-2, scientific(estimate, digits = 3), round(estimate, 2)),
         conf.low = ifelse(`2.5 %` < 10^-2, scientific(`2.5 %`, digits = 3), round(`2.5 %`, 2)),
         conf.high = ifelse(`97.5 %` < 10^-2, scientific(`97.5 %`, digits = 3), round(`97.5 %`, 2))) %>%
  mutate(p.value = scientific(p.value, digits = 3),
         statistic = round(statistic, 2)) %>%
  unite("95% CI", c(conf.low, conf.high), sep = ", ") %>%
  dplyr::select(c(term, estimate, `95% CI`, statistic, p.value)) %>% 
  dplyr::rename("Term" = "term", "Odds ratio" = "estimate",
                "t-statistic" = "statistic", "p-value" = "p.value") %>%
  as.data.frame()

table2c %>% knitr::kable(align = rep('c', 5))

```

#### D. ROC curve of final model

```{r fig3d_prep, eval = FALSE}

## write function to create model on given df

f_log_reg_model <- function(df_model) {

  # resample
  hbox_boot <- bootstraps(df_model, strata = Status, times = 1000)

  # recipe
  hbox_rec <- recipe(Status ~ ., data = df_model) %>% 
    step_BoxCox(all_predictors()) %>% 
    step_nzv(all_predictors()) %>% 
    step_smote(Status) # upsample to deal with unbalanced sample sizes

  # create workflow
  hbox_wf <- workflow() %>% add_recipe(hbox_rec)

  # logistic regression model specifications
  glm_spec <- logistic_reg() %>% set_engine("glm")

  # create model on resamples
  glm_res <- hbox_wf %>% 
    add_model(glm_spec) %>%
    fit_resamples(
      resamples = hbox_boot,
      metrics = metric_set(roc_auc, accuracy, sensitivity, specificity),
      control = tune::control_resamples(save_pred = TRUE, verbose = TRUE)
    )

  # select best model to be used for evaluation
  glm_best <- glm_res %>% select_best("roc_auc")

  # finalize model
  hbox_final <- hbox_wf %>% 
    add_model(glm_spec) %>% 
    finalize_workflow(glm_best) %>%
    fit(df_model) %>%
    pull_workflow_fit()

  # view coefficients & p-values
  model_coefs <- hbox_final %>% tidy()

  # return list including model and a table of best model variable statistics
  list(glm_res, model_coefs)

}

## write function to collect predictions for ROC plot

f_collect_preds <- function(glm_res) {

      glm_res %>% 
      collect_predictions() %>% 
      roc_curve(Status, .pred_CM)

}

## Perform functions on all iterations of model

set.seed(123)

    # select starting variables
    df_hbox_model1 <- df_hbox_log_reg %>% 
      dplyr::select(c(Status, Hematocrit, Lactate, avg_Hb_o2sat, Hb_o2sat_second_alpha, avg_Hb_conc, Hb_conc_second_alpha)) %>% 
      mutate(Status = factor(Status, levels = c("CM", "UM")))

    # model 1: Status ~ Hematocrit + Lactate + avg_Hb_o2sat + Hb_o2sat_second_alpha + avg_Hb_conc + Hb_conc_second_alpha
    model1 <- f_log_reg_model(df_hbox_model1)
    roc1 <- f_collect_preds(model1[[1]]) %>% mutate(model = "1")
    model1[[2]] # remove Hb_o2sat_second_alpha (highest p-val)

    # model 2: Status ~ Hematocrit + Lactate + avg_Hb_o2sat + avg_Hb_conc + Hb_conc_second_alpha
    df_hbox_model2 <-  df_hbox_model1 %>% dplyr::select(-Hb_o2sat_second_alpha)
    model2 <- f_log_reg_model(df_hbox_model2)
    roc2 <- f_collect_preds(model2[[1]]) %>% mutate(model = "2")
    model2[[2]] # remove Lactate

    # model 3: Status ~ Hematocrit + avg_Hb_o2sat + avg_Hb_conc + Hb_conc_second_alpha
    df_hbox_model3 <-  df_hbox_model2 %>% dplyr::select(-Lactate)
    model3 <- f_log_reg_model(df_hbox_model3)
    roc3 <- f_collect_preds(model3[[1]]) %>% mutate(model = "3")
    model3[[2]] # remove avg_Hb_o2sat

    # model 4: Status ~ Hematocrit + avg_Hb_conc + Hb_conc_second_alpha
    df_hbox_model4 <-  df_hbox_model3 %>% dplyr::select(-avg_Hb_o2sat)
    model4 <- f_log_reg_model(df_hbox_model4)
    roc4 <- f_collect_preds(model4[[1]]) %>% mutate(model = "4")
    model4[[2]] # none to remove, final model

save(roc1, roc2, roc3, file = "roc_curves.Rdata")

# calculate areas under curves

library(MESS)
roc1 <- roc1 %>% mutate(false_pos = 1 - specificity)
auc(roc1$false_pos, roc1$sensitivity)

roc2 <- roc2 %>% mutate(false_pos = 1 - specificity)
auc(roc2$false_pos, roc2$sensitivity)

roc3 <- roc3 %>% mutate(false_pos = 1 - specificity)
auc(roc3$false_pos, roc3$sensitivity)

roc_final <- glm_res %>% 
  collect_predictions() %>% 
  roc_curve(Status, .pred_0) %>% 
  mutate(false_pos = 1 - specificity)
auc(roc_final$false_pos, roc_final$sensitivity)

```

```{r fig3d_plot, fig.height = 8}

df_roc <- glm_res %>% 
  collect_predictions() %>% 
  roc_curve(Status, .pred_0) %>% 
  mutate(model = "final") %>% 
  bind_rows(roc1, roc2, roc3)

ggplot() +
  geom_path(data = df_roc %>% filter(model == "final"), 
            aes(x = 1 - specificity, y = sensitivity, color = model), 
            size = 2.5) +
  geom_abline(lty = 3) +
  labs(x = "False positive rate", y = "True positive rate") +
  theme_bw() +

  theme(# text = element_text(family = "Arial"),
        axis.title = element_text(size = 25),
        axis.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 15),
        legend.position = "none")

```

```{r fig3d_inset}

fig3d_inset <- collect_metrics(glm_res) %>% 
  dplyr::select(.metric, mean, std_err) %>% 
  mutate(mean = mean*100, std_err = std_err*100) %>% 
  mutate(Metric = c("Accuracy", "ROC AUC", "Sensitivity", "Specificity")) %>% 
  dplyr::select(c("Metric", "mean", "std_err")) %>%

  dplyr::rename("Mean" = "mean", "Standard error" = "std_err") %>% 
  mutate_if(is.numeric, round, 2) %>% 
  as.data.frame()

fig3d_inset %>% knitr::kable(align = rep('c', 4))

```

### Fig 4. Explanatory model of how to interpret hemoglobin concentration

See manuscript draft (figure created in PowerPoint)

## Supplement
### SFig 1. DFA analysis illustration

#### A. Pre-processed signal
```{r fig_s3a}

df_thc_cumsum1 %>%

  ggplot(aes(x = Time, y = sg_filt)) +
  geom_line() +

  labs(y = "Hemoglobin concentration (μM)", x = "Time (s)") +
  ggtitle("Pre-processed signal") +
  theme_bw() +
  theme(# text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))

```

#### B. Cumulative sum
```{r fig_s3b}

df_thc_cumsum1 %>%

  ggplot(aes(x = Time, y = cumsum)) +
  geom_line() +
  geom_vline(xintercept = 1201, lty = 2, color = "black") +
  geom_vline(xintercept = 1210, lty = 2, color = "black") +

  labs(y = "Cumulative sum", x = "Time (s)") +
  ggtitle("Cumulative signal") +
  theme_bw() +
  theme(# text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))

```

#### C. Window detrend
```{r fig_s3c, fig.show = "hold", out.width = "50%"}

fit <- lm(cumsum ~ Time, data = df_thc_cumsum1 %>% dplyr::filter(Time > 1201 & Time < 1210))

# Window 
df_thc_cumsum1 %>% dplyr::filter(Time > 1201 & Time < 1210) %>%

  ggplot(aes(x = Time, y = cumsum)) +
  geom_line(size = 2) +
  geom_abline(intercept = coef(fit)[1], slope = coef(fit)[2], color = "black", lty = 2, size = 2) +

  labs(y = "Cumulative sum", x = "Time (s)") +
  ggtitle("Cumulative signal") +
  theme_bw() +
  theme(# text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20),
        title = element_text(size = 25))

# Detrended
df_thc_cumsum1 %>% dplyr::filter(Time > 1201 & Time < 1210) %>%

  ggplot(aes(x = Time, y = fit$residuals)) +
  geom_line(size = 2) +
  geom_hline(yintercept = 0, color = "black", lty = 2, size = 2) +

  labs(y = "", x = "Time (s)") +
  ggtitle("Detrended cumulative signal") +
  theme_bw() +
  theme(# text = element_text(family = "Arial"),
        axis.text = element_text(size = 25),
        axis.title = element_text(size = 28),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20),
        title = element_text(size = 25))

```

#### D. Results plot
```{r fig_s3d}

# find slope of segmented line
library(segmented)
df1 <- l_dfa_thc[[1]][[1]] %>% mutate(log10_avg_fluct = log10(avg_fluctuation), log10_window_size = log10(window_size))
fit1 <- lm(log10_avg_fluct ~ log10_window_size, data = df1) 
segfit1 <- segmented(fit1)
slopes1 <- coef(segfit1) # the coefficients are the differences in slope in comparison to the previous slope

# first line: 
#y = b0 + b1*x; y = intercept1 + slope1 * x

# second line:
# y = c0 + c1*x; y = intercept2 + slope2 * x

# At the breakpoint (break1), the segments b and c intersect: b0 + b1*x = c0 + c1*x
b0 <- slopes1[[1]]
b1 <- slopes1[[2]]

c1 <- slopes1[[2]] + slopes1[[3]]
break1 <- segfit1$psi[[2]]

# Solve for c0 (intercept of second segment):
c0 <- b0 + b1 * break1 - c1 * break1

df1 %>%

  ggplot(aes(x = log10_window_size, y = log10_avg_fluct)) +
  geom_point(size = 3, shape = 1) +
  geom_abline(intercept = c0, slope = c1, color = "black", lty = 2) +
  annotate("text", x = -0.5, y = 2, label = paste0("alpha = ", round(c1, 2)), size = 8) +

  labs(x = "log10(window length)", y = "log10(average fluctuation)") +
  ggtitle("DFA results") +

  theme_bw() +
  theme(# text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))

```

### SFig 2. Logistic regression curves

```{r fig_s2, fig.height = 12}

# pivot longer and change factor levels
df_hbox_reg_long <- df_hbox_impute %>% 
  mutate(Status = as.numeric(ifelse(Status == "CM", 1, 0))) %>% 
  pivot_longer(3:ncol(.), names_to = "variable", values_to = "value") %>% 
  mutate(variable = factor(variable, levels = c("Lactate", "Hematocrit", "avg_Hb_o2sat", "avg_Hb_conc", "Hb_o2sat_second_alpha", "Hb_conc_second_alpha")))

# plot as individual regression curves
df_hbox_reg_long %>%

  ggplot(aes(x = value, y = Status)) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_wrap(~variable, scales = "free_x", nrow = 5, ncol = 2) +

  ylab("Status: 1 = CM, 0 = UM") +
  xlab("") +

  theme_bw() +
  theme(# text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20),
        strip.text = element_text(size = 20))

```

**Figure S2.** Plotted are the individual logistic regression curves of individual variables tested for their discriminatory power between UM and CM. On the y-axis, 1 represents CM while 0 represents UM. A steeper curve indicates higher discriminatory power.