### 2023-06-12 Scripts/R/markdown_files/response_to_reviewers.Rmd
12 June 2023 clean and organize scripts for publishing

---
title: "Cerebral oxygenation manuscript response to reviewers"
date: "last updated: `r Sys.Date()`"
output: pdf_document
---

<style type="text/css">
  body{
  font-family: Arial;
  font-size: 12pt;
}
</style>

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "", fig.height = 6, fig.width = 10, 
                      fig.align = 'center')

# Load libraries
library(tidyverse)
library(readxl)
library(DescTools)
library(janitor)
library(FSA)
library(factoextra)
library(ggrepel)
library(ggpubr)
library(tidymodels)
library(vip)
library(scales)
library(reshape)
library(extrafont)
library(grDevices)
library(psych)
library(patchwork)

# Load data
base_dir <- "/Users/ray/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/Ackerman Lab 2019-2021/cerebral_oxygenation/"

# ALT LOADING: use base_dir
#load(paste0(base_dir, "objects/20221125_final_glm_models.Rdata"))
#load(paste0(base_dir, "objects/bandpass_filtered_dfa_results.Rdata"))
#load(paste0(base_dir, "objects/sims.Rdata"))

font_import()

# set plot theme
theme_set(theme_bw() +
              theme(plot.title = element_text(size = 15),
                    axis.title = element_text(size = 15),
                    axis.text = element_text(size = 15),
                    strip.text = element_text(size = 15),
                    legend.title = element_text(size = 15),
                    legend.text = element_text(size = 12)))

```

```{r data}

# read in data

df_hbox <- read_csv(paste0(base_dir, "Data/Hans_datatable_exports/malawi key data v04Jan2022.csv")) %>% 
  as_tibble() %>% 
  janitor::clean_names() %>% 
  mutate(status = replace(status, status == "HV", "HC")) %>% 
  dplyr::rename("subject_id" = "subject_id_session") %>%

  dplyr::filter(status %in% c("CM", "HC", "UM") & session_no == 1 | subject_id == "TM0003CM01") %>% 
  dplyr::filter(!grepl("blood", subject_id)) %>% 
  mutate(status = factor(status, levels = c("HC", "UM", "CM"))) %>% 
  mutate_at(c("bp_dias", "bp_sys", "glucose", "lactate"), as.numeric) %>% # converts remaining character variables to numerics
  select(subject_id, status, everything()) %>% 
  group_by(status)

```

## Reviewer 2

### Minor comment 1

```{r r2c1, fig.height = 10, fig.width = 15}

# MEDIAN (25th, 75th)
df_hbox %>% 
  filter(status == "CM") %>% 
  ungroup %>% 
  select(hrp2, parasitemia) %>% 
  pivot_longer(1:ncol(.), names_to = "variable", values_to = "value") %>% 
  group_by(variable) %>% 
  summarize(median = median(value, na.rm = TRUE),
            `25th` = quantile(value, probs = 0.25, na.rm = TRUE), 
            `75th` = quantile(value, probs = 0.75, na.rm = TRUE)
  ) %>%

  knitr::kable()

# TEST FOR NORMALITY
features <- c("hrp2", "parasitemia", "cerebral_hb_tot", "muscle_hb_tot", "cerebral_hb_tot_alpha2", "muscle_hb_tot_alpha2")

df_normality <- df_hbox %>% 
  filter(status == "CM") %>% 
  ungroup %>% 
  select_at(features) %>% 
  pivot_longer(1:ncol(.), names_to = "feature", values_to = "value") %>%
  group_by(feature) %>% 
  mutate(sw_p = shapiro.test(value)$p.value,
         sw_p = ifelse(sw_p < 0.05, scientific(sw_p, digits = 3), as.character(round(sw_p, 2))),
         sw_p = paste0("SW p = ", sw_p),

         xmin = density(value, na.rm = TRUE)$x %>% min %>% floor,
         xmax = density(value, na.rm = TRUE)$x %>% max %>% ceiling,
         ymin = density(value, na.rm = TRUE)$y %>% min %>% floor,
         ymax = density(value, na.rm = TRUE)$y %>% max %>% ceiling

  )

df_text <- df_normality %>% 
  dplyr::select(-value) %>% 
  distinct() %>% 
  mutate(x = xmax - 0.1*xmax,
         y = ymax - 0.1*ymax) %>% 
  dplyr::select(-contains("min"), -contains("max")) %>% 
  dplyr::rename("label" = "sw_p")

p_sw <- df_normality %>% 
  ggplot(aes(x = value)) +
  geom_density(trim = FALSE) +
  geom_text(
    data = df_text,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1,
    vjust = 1,
    size = 5
  ) + 
  labs(y = "") +
  facet_wrap(vars(feature), scales = "free")

# PLOT CORRELATIONS BETWEEN QUANTITATIVE PARASITEMIA MEASUREMENTS IN CM AND CEREBRAL HBTOT
p_hrp2_cerebral <- df_hbox %>% 
  filter(status == "CM") %>%

  ggplot(aes(x = log10(hrp2), y = log10(cerebral_hb_tot))) +
  geom_point(size = 4, shape = 1) +
  geom_smooth(method = "lm") +
  stat_cor(size = 5) +
  labs(x = paste0("log10(HRP2) (ng/\u03bcL)"), y = "log10(cerebral Hbtot)") +
  ggtitle("NIRS-derived cerebral Hbtot and HRP2 in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

p_parasitemia_cerebral <- df_hbox %>% 
  filter(status == "CM") %>%

  ggplot(aes(x = log10(parasitemia), y = log10(cerebral_hb_tot))) +
  geom_point(size = 4, shape = 1) +
  geom_smooth(method = "lm") +
  stat_cor(size = 5) +
  labs(x = "log10(parasitemia) (RBCs/\u03bcL)", y = "log10(cerebral Hbtot)") +
  ggtitle("NIRS-derived cerebral Hbtot and parasitemia in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

# PLOT CORRELATIONS BETWEEN QUANTITATIVE PARASITEMIA MEASUREMENTS IN CM AND MUSCle HBTOT
p_hrp2_muscle <- df_hbox %>% 
  filter(status == "CM") %>%

  ggplot(aes(x = log10(hrp2), y = muscle_hb_tot)) +
  geom_point(size = 4, shape = 1) +
  geom_smooth(method = "lm") +
  stat_cor(size = 5) +
  xlim(c(2.5, 4.5)) +
  labs(x = paste0("log10(HRP2) (ng/\u03bcL)"), y = bquote(`Muscle Hb`[tot])) +
  ggtitle("NIRS-derived muscle Hbtot and HRP2 in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

p_parasitemia_muscle <- df_hbox %>% 
  filter(status == "CM") %>%

  ggplot(aes(x = log10(parasitemia), y = muscle_hb_tot)) +
  geom_point(size = 4, shape = 1) +
  geom_smooth(method = "lm") +
  stat_cor(size = 5) +
  labs(x = "log10(parasitemia) (RBCs/\u03bcL)", y = bquote(`muscle Hb`[tot])) +
  ggtitle("NIRS-derived muscle Hbtot and parasitemia in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

# PLOT CORRELATIONS BETWEEN QUANTITATIVE PARASITEMIA MEASUREMENTS IN CM AND CEREBRAL HBTOT ALPHA
p_hrp2_cerebral_alpha <- df_hbox %>% 
  filter(status == "CM") %>%

  ggplot(aes(x = log10(hrp2), y = cerebral_hb_tot_alpha2)) +
  geom_point(size = 4, shape = 1) +
  geom_smooth(method = "lm") +
  stat_cor(size = 5) +
  labs(x = paste0("log10(HRP2) (ng/\u03bcL)"), y = "Cerebral Hbtot alpha") +
  ggtitle("NIRS-derived cerebral Hbtot alpha and HRP2 in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

p_parasitemia_cerebral_alpha <- df_hbox %>% 
  filter(status == "CM") %>%

  ggplot(aes(x = log10(parasitemia), y = cerebral_hb_tot_alpha2)) +
  geom_point(size = 4, shape = 1) +
  geom_smooth(method = "lm") +
  stat_cor(size = 5) +
  labs(x = "log10(parasitemia) (RBCs/\u03bcL)", y = "Cerebral Hbtot alpha") +
  ggtitle("NIRS-derived cerebral Hbtot alpha and parasitemia in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

# PLOT CORRELATIONS BETWEEN QUANTITATIVE PARASITEMIA MEASUREMENTS IN CM AND MUSCLE HBTOT ALPHA
p_hrp2_muscle_alpha <- df_hbox %>% 
  filter(status == "CM") %>%

  ggplot(aes(x = log10(hrp2), y = muscle_hb_tot_alpha)) +
  geom_point(size = 4, shape = 1) +
  geom_smooth(method = "lm") +
  stat_cor(size = 5) +
  xlim(c(2.5, 4.5)) +
  labs(x = paste0("log10(HRP2) (ng/\u03bcL)"), y = "Muscle Hbtot alpha") +
  ggtitle("NIRS-derived muscle Hbtot alpha and HRP2 in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

p_parasitemia_muscle_alpha <- df_hbox %>% 
  filter(status == "CM") %>%

  ggplot(aes(x = log10(parasitemia), y = muscle_hb_tot_alpha2)) +
  geom_point(size = 4, shape = 1) +
  geom_smooth(method = "lm") +
  stat_cor(size = 5) +
  labs(x = "log10(parasitemia) (RBCs/\u03bcL)", y = "Muscle Hbtot alpha") +
  ggtitle("NIRS-derived muscle Hbtot alpha and parasitemia in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

# COMBINE PLOTS

p_sw

(p_hrp2_cerebral | p_parasitemia_cerebral) /
(p_hrp2_muscle | p_parasitemia_muscle)

(p_hrp2_cerebral_alpha | p_parasitemia_cerebral_alpha) /
(p_hrp2_muscle_alpha | p_parasitemia_muscle_alpha)

```

## Reviewer 3

### Recommendation 1

```{r r3r1}

# REGRESSION
# df_hbox %>% 
#   filter(status == "CM") %>% 
#   
#   ggplot(aes(x = cerebral_hb_tot, y = outcome)) +
#   geom_point(size = 4, shape = 1) +
#   geom_smooth(method = "lm") +
#   stat_cor(size = 5) +
#   labs(x = "Cerebral Hbtot", y = "Outcome") +
#   ggtitle("Correlation between NIRS-derived cerebral Hbtot and outcome in children with CM") +
#   theme_bw() +
#   theme(axis.text = element_text(size = 15),
#         axis.title = element_text(size = 15))

# KRUSKAL-WALLIS
kw_p <- kruskal.test(cerebral_hb_tot ~ outcome,
             data = df_hbox %>%
               filter(status == "CM") %>% 
               mutate(outcome = as.factor(outcome))) %>% 
  .$p.value %>% 
  round(2)

df_hbox %>%
  filter(status == "CM") %>%
  mutate(outcome =

           case_when(

             outcome == 1 ~ "survival",
             outcome == 2 ~ "survival with neurological sequelae",
             outcome == 3 ~ "non-survival"

           ) %>%

           factor(levels = c("survival", "survival with neurological sequelae", "non-survival"))

           ) %>%

  ggplot(aes(x = outcome, y = cerebral_hb_tot, color = outcome)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(size = 4, alpha = 0.5,
             position = position_jitter(width = 0.1)) +
  #annotate(geom = "text", label = paste0("KW p-value = ", kw_p), x = 3, y = 350, size = 5) +
  labs(x = "", y = bquote(`Cerebral Hb`[tot])) +
  ggtitle("Outcome and NIRS-derived cerebral Hbtot in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.position = "none"
        )

```