### 2023-06-12 Scripts/R/markdown_files/20221205_brain_swell_model.Rmd

---
output: pdf_document
---

Characterize the relationship between brain swelling and cerebral Hb_tot

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.height = 5.5, fig.width = 8)

# libraries
library(tidyverse)
library(RColorBrewer)
library(gridExtra)
library(patchwork)
library(Hmisc)
library(corrr)
library(tidymodels)
library(themis)
library(vip)
library(janitor)
library(kableExtra)

# set theme for plots
theme_set(theme_bw() +
            theme(plot.title = element_text(size = 18),
                  axis.title = element_text(size = 15),
                  axis.text = element_text(size = 12),
                  strip.text = element_text(size = 12),
                  legend.title = element_text(size = 15),
                  legend.text = element_text(size = 12)))

# create color vector
my_col <- c("CM" = "#00BFC4", "HC" = "#7CAE00", "UM" = "#F8766D")

```

```{r data, echo = FALSE}

# load
df_hbox <- read_csv("Data/Hans_datatable_exports/malawi key data v04Jan2022.csv") %>% 
  as_tibble() %>% 
  janitor::clean_names() %>% 
  mutate(status = replace(status, status == "HV", "HC")) %>% 
  dplyr::rename("subject_id" = "subject_id_session") %>%

  dplyr::filter(status %in% c("CM", "HC", "UM") & session_no == 1 | subject_id == "TM0003CM01") %>% 
  dplyr::filter(!grepl("blood", subject_id)) %>% 
  mutate(status = factor(status, levels = c("HC", "UM", "CM"))) %>% 
  mutate_at(c("bp_dias", "bp_sys", "glucose", "lactate"), as.numeric) %>% # converts remaining character variables to numerics
  select(subject_id, status, everything()) %>% 
  group_by(status) %>% 
  mutate(log10_cerebral_hb_tot = log10(cerebral_hb_tot))

```

```{r fig1}

# extract hb_tot and hb_oxy for the brain and muscle
df_fig1 <- df_hbox %>% 
  select(contains(c("hb_tot","hb_oxy")), -contains(c("alpha", "bsl"))) %>%
  ungroup() %>% 
  pivot_longer(2:ncol(.), names_to = "variable", values_to = "value") %>% 
  separate(variable, into = c("tissue", "variable"), sep = "_", extra = "merge")

# create color vector
my_col <- c("CM" = "#00BFC4", "HC" = "#7CAE00", "UM" = "#F8766D")

# plot
df_fig1 %>%

  ggplot(aes(x = status, y = value)) +
  geom_violin(aes(color = status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1, size = 2) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = status), size = 0.2, width = 0.5) +
  facet_grid(variable ~ tissue, scales = "free_y") +

  scale_color_manual(values = my_col) +
  labs(x = "") +
  theme_bw() +
  theme(legend.position = "none")  +

  theme(# text = element_text(family = "Arial"),
        strip.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        axis.title.y = element_blank(),
        axis.text = element_text(size = 15))

```

## Brain swell model fitting

1. Log10-transformed Hb_tot vs brain swelling – with hematocrit indicated by plotting character size (with all data points represented)
2. A simple plot of hct vs brain swelling for my own understanding of that relationship.

```{r plot}

df_hbox %>%
  ggplot(aes(x = log10_cerebral_hb_tot, y = brain_swell)) +
  geom_point(aes(size = hct), shape = 1) +

  geom_smooth(method = "lm", lty = 2, color = "#696969", alpha = 0.2, se = FALSE) +

  theme_bw() +
  ggtitle("1. Relationship between brain swell score and log10-transformed cerebral Hb_tot") +
  theme(strip.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        axis.text = element_text(size = 15),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))

df_hbox %>%
  ggplot(aes(x = hct, y = brain_swell)) +
  geom_point(shape = 1, size = 2) +

  geom_smooth(method = "lm", lty = 2, color = "#696969", alpha = 0.2, se = FALSE) +
  theme_bw() +
  ggtitle("2. Relationship between hematocrit & brain swell score") +
  theme(strip.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        axis.text = element_text(size = 15),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))

```

<!-- ```{r brain_swell} -->

<!-- cor.test( ~ brain_swell + log10_cerebral_hb_tot,  -->
<!--           data = df_hbox, -->
<!--           method = "spearman") -->

<!-- lm(brain_swell ~ log10_cerebral_hb_tot, data = df_hbox) %>% summary -->

<!-- lm(brain_swell ~ log10_cerebral_hb_tot + hct, data = df_hbox) %>% summary -->

<!-- lm(brain_swell ~ log10_cerebral_hb_tot + cerebral_hb_tot_alpha2, data = df_hbox) %>% summary -->

<!-- lm(brain_swell ~ log10_cerebral_hb_tot + cerebral_hb_tot_alpha2 + hct, data = df_hbox) %>% summary -->

<!-- lm(brain_swell ~ log10_cerebral_hb_tot + cerebral_hb_tot_alpha2, data = df_hbox) %>% summary -->

<!-- lm(brain_swell ~ cerebral_hb_oxy + cerebral_hb_oxy_alpha2 + log10_cerebral_hb_tot, data = df_hbox) %>% summary -->

<!-- ``` -->