### 2023-06-12 Scripts/R/markdown_files/20221217_parasitemia.Rmd

---
title: "parasitemia"
date: "`r Sys.Date()`"
output: pdf_document
---

Identify differences in parastiemia counts between patient groups

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.height = 5.5, fig.width = 8)

# libraries
library(tidyverse)
library(RColorBrewer)
library(patchwork)
library(Hmisc)
library(corrr)
library(tidymodels)
library(themis)
library(vip)
library(janitor)
library(kableExtra)
library(MASS)

# set theme for plots
theme_set(theme_bw() +
            theme(plot.title = element_text(size = 18),
                  axis.title = element_text(size = 15),
                  axis.text = element_text(size = 12),
                  strip.text = element_text(size = 12),
                  legend.title = element_text(size = 15),
                  legend.text = element_text(size = 12)))

# create color vector
my_col <- c("CM" = "#00BFC4", "HC" = "#7CAE00", "UM" = "#F8766D")

```

```{r data, echo = FALSE}

# load
df_hbox <- read_csv("Data/Hans_datatable_exports/malawi key data v04Jan2022.csv") %>% 
  as_tibble() %>% 
  janitor::clean_names() %>% 
  mutate(status = replace(status, status == "HV", "HC")) %>% 
  dplyr::rename("subject_id" = "subject_id_session") %>%

  dplyr::filter(status %in% c("CM", "HC", "UM") & session_no == 1 | subject_id == "TM0003CM01") %>% 
  dplyr::filter(!grepl("blood", subject_id)) %>% 
  mutate(status = factor(status, levels = c("HC", "UM", "CM"))) %>% 
  mutate_at(c("bp_dias", "bp_sys", "glucose", "lactate"), as.numeric) %>% # converts remaining character variables to numerics
  dplyr::select(subject_id, status, everything()) %>% 
  group_by(status)

# format
  # define clinical variables of interest
  clinical_voi <- c("age_calc", "temperature", "hr", "bp_sys", "bp_dias", "resp_rate", "pulse_ox", "hct", "lactate", "glucose", "aa_arg")
  hbox_voi <- c("cerebral_hb_tot", "cerebral_hb_tot_alpha2")

  # select clinical VOI
  df_model <- df_hbox %>% 
    dplyr::select(c(status, sex, all_of(hbox_voi), all_of(clinical_voi))) %>% 
    mutate(sex_coded = ifelse(sex == "Female", 1, 0)) %>% 
    dplyr::select(-sex)

  # impute missing data with median by group
  f_impute <- function(x, na.rm = TRUE) (replace(x, is.na(x), median(x, na.rm = na.rm)))

  df_model_impute <- df_model %>% 
    group_by(status) %>% 
    mutate_at(2:ncol(.), f_impute) %>% 
    ungroup()

```

## Parasite counts in each patient group
```{r parasite_counts}

df_parasite_counts <- df_hbox %>% 
  dplyr::select(parasites) %>% 
  dplyr::count(parasites) %>% 
  pivot_wider(id_cols = parasites, names_from = status, values_from = n) %>% 
  dplyr::filter(!is.na(parasites)) %>% # 1 CM patient had no value for parasites
  dplyr::mutate_if(is.numeric, ~ ifelse(is.na(.x), 0, .x))

df_parasite_counts %>% knitr::kable(align = rep('c', 4))

# plot
df_parasite_counts %>% 
  pivot_longer(2:ncol(.), names_to = "status", values_to = "count") %>% 
  dplyr::mutate(status = factor(status, levels = c("HC", "UM", "CM")),
         parasites = factor(parasites, levels = c("Neg", "1+", "2+", "3+", "4+", "5+")),
         count = ifelse(count == 0, NA_real_, count)) %>%

  ggplot(aes(x = parasites, y = count, fill = status)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = my_col) +
  ylab("") +
  ggtitle("Parasite counts in each patient group")

```

## Statistical tests to compare parasites in each patient group

### Chi-squared
```{r chisq}

# Chi-squared of UM vs CM
chisq.test(df_parasite_counts %>% 
             dplyr::select(-HC) %>% 
             as.data.frame %>% 
             column_to_rownames("parasites")
)

```

### Ordinal regression (?)
```{r ord}

df_parasites <- df_hbox %>% 
  dplyr::select(parasites) %>% 
  mutate(parasites = factor(parasites, levels = c("Neg", "1+", "2+", "3+", "4+", "5+")))

ord <- polr(status ~ parasites, data = df_parasites, Hess = TRUE)
tidy(ord) %>% 
  mutate(p_val = pnorm(abs(statistic), lower.tail = FALSE)*2) %>% 
  clean_names %>% 
  dplyr::rename("coefficient" = "estimate",
                "t_stat" = "statistic")

```

### Chi-squared (binarize parasites variable)
```{r chisq_bin}

df_parasite_counts_binned <- df_parasite_counts %>% 
  mutate(parasites_bin = ifelse(parasites %in% c("Neg", "1+", "2+", "3+"),
                                "low", "high") %>% 
           factor(levels = c("low", "high"))) %>% 
  dplyr::select(parasites_bin, UM, CM) %>% 
  group_by(parasites_bin) %>% 
  summarise(UM = sum(UM), CM = sum(CM))

df_parasite_counts_binned %>% knitr::kable(align = rep('c', 3))

# Chi-squared of UM vs CM
chisq.test(df_parasite_counts_binned %>% 
             as.data.frame %>% 
             column_to_rownames("parasites_bin")
)

```

### Logistic regression (binarize parasites variable)
```{r logreg}

df_parasites_binned <- df_parasites %>% 
  mutate(parasites_bin = ifelse(parasites %in% c("Neg", "1+", "2+", "3+"),
                                "low", "high") %>% 
           factor(levels = c("low", "high")))

# Chi-squared of UM vs CM
logreg <- glm(status ~ parasites_bin, data = df_parasites_binned, family = "binomial")
tidy(logreg) %>% 
  clean_names %>% 
  dplyr::rename("coefficient" = "estimate",
                "t_stat" = "statistic")

```

### HRP & parasitemia in CM patients
```{r parasitemia}

cm_voi <- c("hrp2", "parasitemia")

df_cm_voi <- df_hbox %>% 
  ungroup %>% 
  filter(status == "CM") %>% 
  dplyr::select_at(cm_voi)

# HRP
df_cm_hrp2_parasitemia <- df_cm_voi %>% 
  summarise_all(funs(median = median, 
                     q1 = quantile(., probs = 0.25), 
                     q3 = quantile(., probs = 0.75)), 
                na.rm = TRUE) %>% 
  mutate(hrp2 = paste0(hrp2_median, " (", hrp2_q1, ", ", hrp2_q3, ")"),
         parasitemia = paste0(parasitemia_median, " (", parasitemia_q1, ", ", parasitemia_q3, ")")) %>% 
  dplyr::select(hrp2, parasitemia)

df_cm_hrp2_parasitemia %>% knitr::kable(align = rep('c', 2))

```