### 2023-06-12 Scripts/R/markdown_files/archive/fig3a_ideas.Rmd

---
title: "fig3a"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "", fig.height = 6, fig.width = 10)

# libraries
library(tidyverse)
library(psych)
library(stringr)

```

```{r data}

# read in data

df_hbox <- read_csv("Data/Hans_datatable_exports/malawi key data v04Jan2022.csv") %>% 
  as_tibble() %>% 
  janitor::clean_names() %>% 
  mutate(status = replace(status, status == "HV", "HC")) %>% 
  dplyr::rename("subject_id" = "subject_id_session") %>%

  dplyr::filter(status %in% c("CM", "HC", "UM") & session_no == 1 | subject_id == "TM0003CM01") %>% 
  dplyr::filter(!grepl("blood", subject_id)) %>% 
  mutate(status = factor(status, levels = c("HC", "UM", "CM"))) %>% 
  mutate_at(c("bp_dias", "bp_sys", "glucose", "lactate"), as.numeric) %>% # converts remaining character variables to numerics
  select(subject_id, status, everything()) %>% 
  group_by(status)

```

```{r fig3_data}

# select model variables of interest
model_voi <- c("subject_id", "status", 
               "hct", "lactate",
               "cerebral_hb_tot", "cerebral_hb_oxy",
               "cerebral_hb_tot_alpha2", "cerebral_hb_oxy_alpha2")

df_model <- df_hbox %>% 
  select_at(model_voi) %>%

  # remove HC
  ungroup %>% 
  filter(status != "HC")

# impute missing data with median by group
  f_impute <- function(x, na.rm = TRUE) (replace(x, is.na(x), median(x, na.rm = na.rm)))

  df_model_impute <- df_model %>% 
    group_by(status) %>% 
    mutate_at(3:ncol(.), f_impute) %>% 
    ungroup()

```

```{r fig3a_cors}

# find pearson correlation across all variables of interest, including p-values and adjusted p-values
cm_cor <- df_model %>% 
  dplyr::filter(status == "CM") %>% 
  dplyr::select(-c(subject_id, status)) %>% 
  corr.test(method = "pearson", adjust = "none")

um_cor <- df_model %>% 
  dplyr::filter(status == "UM") %>% 
  dplyr::select(-c(subject_id, status)) %>% 
  corr.test(method = "pearson", adjust = "none")

# combine into tibble
df_fig3a <- cm_cor$r %>% 
  as_tibble %>% 
  mutate(var1 = colnames(.), .before = hct) %>% 
  pivot_longer(2:ncol(.), names_to = "var2", values_to = "pearsons_r") %>% 
  mutate(status = "CM", .before = var1) %>%

  left_join(

    cm_cor$p %>% 
      as_tibble %>% 
      mutate(var1 = colnames(.), .before = hct) %>% 
      pivot_longer(2:ncol(.), names_to = "var2", values_to = "p_value"),

    by = c("var1", "var2")

  ) %>%

  # adjust p-values
  mutate(adj_p_val = p.adjust(p_value, method = "BH")) %>%

  bind_rows(

    um_cor$r %>% 
      as_tibble %>% 
      mutate(var1 = colnames(.), .before = hct) %>% 
      pivot_longer(2:ncol(.), names_to = "var2", values_to = "pearsons_r") %>% 
      mutate(status = "UM", .before = var1) %>%

      left_join(

        um_cor$p %>% 
          as_tibble %>% 
          mutate(var1 = colnames(.), .before = hct) %>% 
          pivot_longer(2:ncol(.), names_to = "var2", values_to = "p_value"),

        by = c("var1", "var2")

      ) %>% 
      mutate(adj_p_val = p.adjust(p_value, method = "BH"))

  ) %>%

  # set factor levels for variables
  mutate_at(2:3, factor, levels = c("hct", "lactate", "cerebral_hb_oxy", "cerebral_hb_tot", "cerebral_hb_oxy_alpha2", "cerebral_hb_tot_alpha2")) %>%

  # set diagonal equal to zero
  mutate(pearsons_r = ifelse(var1 == var2, 0, pearsons_r))

```

### idea 1: put box around values significant by p-value
```{r fig3a_v1}

df_fig3a %>% 
  mutate(is_sig = ifelse(p_value < 0.05 & var1 != var2, "yes", "no")) %>%

  ggplot(aes(x = var1, y = var2, fill = pearsons_r, label = ifelse(pearsons_r != 0, round(pearsons_r, 2), NA))) + 
  geom_tile(aes(width = 0.9, height = 0.9)) +
  geom_tile(aes(color = is_sig, width = 0.9, height = 0.9), size = 1, show.legend = FALSE) +

  facet_wrap(~ status) +

  scale_color_manual(values = c("white", "black")) +
  scale_fill_gradient2(low = "#4575b4", high = "#d73027") +
  geom_text(size = 4) +

  xlab("") +
  ylab("") +
  theme_classic() +
  theme(axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 0.95))

```

### idea 2: put box around values significant by BH-adjusted p-value
```{r fig3a_v2}

df_fig3a %>% 
  mutate(is_sig = ifelse(adj_p_val < 0.05 & var1 != var2, "yes", "no")) %>%

  ggplot(aes(x = var1, y = var2, fill = pearsons_r, label = ifelse(pearsons_r != 0, round(pearsons_r, 2), NA))) + 
  geom_tile(aes(width = 0.9, height = 0.9)) +
  geom_tile(aes(color = is_sig, width = 0.9, height = 0.9), size = 1, show.legend = FALSE) +

  facet_wrap(~ status) +

  scale_color_manual(values = c("white", "black")) +
  scale_fill_gradient2(low = "#4575b4", high = "#d73027") +
  geom_text(size = 4) +

  xlab("") +
  ylab("") +
  theme_classic() +
  theme(axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 0.95))

```

### idea 3: add p-value to box label
```{r fig3a_v3}

df_fig3a %>% 
  mutate(label = paste0(round(pearsons_r, 2), " ", round(p_value, 2))) %>% 
  mutate(label = ifelse(var1 == var2, NA, label)) %>% 
  mutate(label = str_wrap(label, width = 2)) %>%

  ggplot(aes(x = var1, y = var2, fill = pearsons_r, label = label)) + 
  geom_tile(aes(width = 0.95, height = 0.95)) +

  facet_wrap(~ status) +

  scale_fill_gradient2(low = "#4575b4", high = "#d73027") +
  geom_text(size = 3) +

  xlab("") +
  ylab("") +
  theme_classic() +
  theme(axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 0.95))

```

### idea 3: add BH-adjusted p-value to box label
```{r fig3a_v4}

df_fig3a %>% 
  mutate(label = paste0(round(pearsons_r, 2), " ", round(adj_p_val, 2))) %>% 
  mutate(label = ifelse(var1 == var2, NA, label)) %>% 
  mutate(label = str_wrap(label, width = 2)) %>%

  ggplot(aes(x = var1, y = var2, fill = pearsons_r, label = label)) + 
  geom_tile(aes(width = 0.95, height = 0.95)) +

  facet_wrap(~ status) +

  scale_fill_gradient2(low = "#4575b4", high = "#d73027") +
  geom_text(size = 3) +

  xlab("") +
  ylab("") +
  theme_classic() +
  theme(axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 0.95))

```