### 2023-06-12 Scripts/R/markdown_files/archive/20221117_revise_model.Rmd
add objects to run scripts, other data will be available in data repository

---
title: "Revise LR model"
author: "Rachel Smith"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.height = 5.5, fig.width = 8)

# libraries
library(tidyverse)
library(RColorBrewer)
library(gridExtra)
library(patchwork)
library(Hmisc)
library(corrr)
library(tidymodels)
library(themis)
library(vip)
library(janitor)

# set theme for plots
theme_set(theme_bw() +
            theme(plot.title = element_text(size = 18),
                  axis.title = element_text(size = 15),
                  axis.text = element_text(size = 12),
                  strip.text = element_text(size = 12),
                  legend.title = element_text(size = 15),
                  legend.text = element_text(size = 12)))

# create color vector
my_col <- c("CM" = "#00BFC4", "HC" = "#7CAE00", "UM" = "#F8766D")

```

```{r data, echo = FALSE}

# load
df_hbox <- read_csv("Data/Hans_datatable_exports/malawi key data v04Jan2022.csv") %>% 
  as_tibble() %>% 
  janitor::clean_names() %>% 
  mutate(status = replace(status, status == "HV", "HC")) %>% 
  dplyr::rename("subject_id" = "subject_id_session") %>%

  dplyr::filter(status %in% c("CM", "HC", "UM") & session_no == 1 | subject_id == "TM0003CM01") %>% 
  dplyr::filter(!grepl("blood", subject_id)) %>% 
  mutate(status = factor(status, levels = c("HC", "UM", "CM"))) %>% 
  mutate_at(c("bp_dias", "bp_sys", "glucose", "lactate"), as.numeric) %>% # converts remaining character variables to numerics
  select(subject_id, status, everything()) %>% 
  group_by(status)

# format
  # define clinical variables of interest
  clinical_voi <- c("age_calc", "temperature", "hr", "bp_sys", "bp_dias", "resp_rate", "pulse_ox", "hct", "lactate", "glucose", "aa_arg")
  hbox_voi <- c("cerebral_hb_tot", "cerebral_hb_tot_alpha2")

  # select clinical VOI
  df_model <- df_hbox %>% 
    dplyr::select(c(status, sex, all_of(hbox_voi), all_of(clinical_voi))) %>% 
    mutate(sex_coded = ifelse(sex == "Female", 1, 0)) %>% 
    dplyr::select(-sex)

  # impute missing data with median by group
  f_impute <- function(x, na.rm = TRUE) (replace(x, is.na(x), median(x, na.rm = na.rm)))

  df_model_impute <- df_model %>% 
    group_by(status) %>% 
    mutate_at(2:ncol(.), f_impute) %>% 
    ungroup()

```

```{r p_distributions, fig.height = 8, fig.width = 10}

df_model %>% 
  pivot_longer(2:ncol(.), names_to = "variable", values_to = "value") %>% 
  mutate(variable = factor(variable, levels = c(clinical_voi, "sex_coded", hbox_voi))) %>% 
  filter(!(variable == "age_calc" & value > 20000)) %>%

  ggplot(aes(x = value)) +
  geom_density(aes(fill = status), alpha = 0.5) +
  facet_wrap(~ variable, scales = "free") +
  labs(x = "") +
  ggtitle("Distributions of variables of interest")

```

```{r correlate}

# correlation
df_cor_r <- df_model_impute %>% 
  group_by(status) %>% 
  nest() %>% 
  mutate(corr = map(.x = data, 
                    .f = ~ .x %>% 
                      correlate %>% 
                      dplyr::select(term, contains("cerebral_hb_tot")) %>% 
                      filter(term %in% clinical_voi) 
  )
  ) %>% 
  unnest(cols = c(corr)) %>% 
  pivot_longer(4:5, names_to = "voi", values_to = "r") %>% 
  mutate(voi = ifelse(voi == "cerebral_hb_tot", "Hb_tot", "Hb_tot_a") %>% factor()) %>% 
  mutate(term = factor(term, levels = clinical_voi))

# p-values
df_cor_p <- df_model_impute %>% 
  group_by(status) %>% 
  nest() %>% 
  mutate(corr = map(.x = data, 
                    .f = ~ .x %>% 
                      as.data.frame %>% 
                      as.matrix %>% 
                      rcorr %>% 
                      .$P %>% 
                      corrr::as_cordf() %>% 
                      dplyr::select(term, contains("cerebral_hb_tot")) %>% 
                      filter(term %in% clinical_voi) 
  )
  ) %>% 
  unnest(cols = c(corr)) %>% 
  pivot_longer(4:5, names_to = "voi", values_to = "p") %>% 
  mutate(voi = ifelse(voi == "cerebral_hb_tot", "Hb_tot", "Hb_tot_a") %>% factor()) %>% 
  mutate(term = factor(term, levels = clinical_voi))

# combine for plot
df_plot_cor <- df_cor_r %>% 
  left_join(df_cor_p) %>% #filter(status == "HC")
  mutate(label = as.character(round(r, 2)),
         label = ifelse(p < 0.05, paste0(label, "*"), label))

```

```{r p_correlations}

df_plot_cor %>% 
  ggplot() +
  geom_tile(aes(x = voi, y = term, fill = r)) +
  geom_text(aes(x = voi,y = term, label = label)) +
  facet_wrap( ~ status) +
  labs(x = "", y = "") +
  scale_fill_gradient2(high = "#B2182B", mid = "white", low = "#2166AC",
                       limits = c(-1, 1)) +
  ylim(rev(levels(df_cor_p$term))) +
  ggtitle("Correlations of clinical variables with cerebral Hb_tot") +
  theme_classic() +
  theme(plot.title = element_text(size = 18),
        axis.title = element_text(size = 15),
        axis.text = element_text(size = 12),
        strip.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        legend.text = element_text(size = 12))

```
*denotes significant correlation (p < 0.05)

```{r p_sig_cor}

p_um1 <- df_hbox %>% 
  filter(status == "UM") %>% 
  dplyr::select(cerebral_hb_tot, hr) %>%

  ggplot(aes(x = cerebral_hb_tot, y = hr, color = status)) +
  geom_point(size = 2, show.legend = FALSE) +
  geom_smooth(method = "lm", color = "black", lty = 2, se = FALSE) +
  annotate(geom = "text", 
           label = "r = -0.44; p < 0.01",
           x = 300, y = 160,
           size = 5) + 
  scale_color_manual(values = my_col) +
  ggtitle("UM HR & Hb_tot")

p_um2 <- df_hbox %>% 
  filter(status == "UM") %>% 
  dplyr::select(cerebral_hb_tot, hct) %>%

  ggplot(aes(x = cerebral_hb_tot, y = hct, color = status)) +
  geom_point(size = 2, show.legend = FALSE) +
  geom_smooth(method = "lm", color = "black", lty = 2, se = FALSE) +
  annotate(geom = "text",
           label = "r = 0.39; p = 0.03",
           x = 300, y = 27,
           size = 5) +
  scale_color_manual(values = my_col) +
  ggtitle("UM Hematocrit & Hb_tot")

p_hc1 <- df_hbox %>% 
  filter(status == "HC") %>% 
  dplyr::select(cerebral_hb_tot, aa_arg) %>%

  ggplot(aes(x = cerebral_hb_tot, y = aa_arg, color = status)) +
  geom_point(size = 2, show.legend = FALSE) +
  geom_smooth(method = "lm", color = "black", lty = 2, se = FALSE) +
  annotate(geom = "text",
           label = "r = 0.51; p < 0.01",
           x = 95, y = 45,
           size = 5) +
  scale_color_manual(values = my_col) +
  ggtitle("HC arg & Hb_tot")

p_hc2 <- df_hbox %>% 
  filter(status == "HC") %>% 
  dplyr::select(cerebral_hb_tot_alpha, aa_arg) %>%

  ggplot(aes(x = cerebral_hb_tot_alpha, y = aa_arg, color = status)) +
  geom_point(size = 2, show.legend = FALSE) +
  geom_smooth(method = "lm", color = "black", lty = 2, se = FALSE) +
  annotate(geom = "text",
           label = "r = -0.43; p = 0.02",
           x = 1.02, y = 45,
           size = 5) +
  scale_color_manual(values = my_col) +
  ggtitle("HC arg & Hb_tot_a")

(p_um1 + p_um2) / (p_hc1 + p_hc2)

```

```{r lr_model}

# write function to run LR model
f_run_lr <- function(df_model, resamples) {

  # resample
  hbox_boot <- bootstraps(df_model, strata = status, times = resamples)

  # recipe
  hbox_rec <- recipe(status ~ ., data = df_model) %>% 
    step_impute_knn(all_predictors()) %>% 
    step_normalize(all_predictors()) %>%
    step_zv(all_predictors()) %>%
    step_smote(status) %>% 
    update_role(subject_id, new_role = "id")

  # create workflow
  hbox_wf <- workflow() %>% add_recipe(hbox_rec)

  # logistic regression model specifications
  glm_spec <- logistic_reg() %>% set_engine("glm")

  # create model on 5000 different resamples
  glm_res <- hbox_wf %>% 
    add_model(glm_spec) %>%
    fit_resamples(
      resamples = hbox_boot,
      metrics = metric_set(roc_auc, accuracy, sensitivity, specificity),
      control = tune::control_resamples(save_pred = TRUE, verbose = TRUE)
    )

  # select best model to be used for evaluation
  glm_best <- glm_res %>% select_best("roc_auc")

  # finalize model
  hbox_wf %>% 
    add_model(glm_spec) %>% 
    finalize_workflow(glm_best) %>%
    fit(df_model) %>%
    extract_fit_parsnip()

}

```

### Cerebral Hb_tot model fitting

status ~ Hb_tot + age_calc + bp_sys + bp_dias + hr + hct + lactate + aa_arg + sex
```{r hb_tot_model1}

# VOI: age_calc, bp_sys, bp_dias, hr, hct, lactate, and aa_arg plus either Hb_tot or Hb_tot_a. I would probably include sex as well

df_model <- df_hbox %>% 
  mutate(sex = ifelse(sex == "Female", 1, 0)) %>% 
  filter(status != "HC") %>% 
  mutate(status = factor(status, levels = c("UM", "CM")))

set.seed(20221117)

# 1. status ~ Hb_tot + age_calc + bp_sys + bp_dias + hr + hct + lactate + aa_arg + sex
hbtot_final1 <- f_run_lr(df_model %>% 
                           dplyr::select(subject_id, status, cerebral_hb_tot, 
                                         aa_arg, age_calc, bp_sys, bp_dias, hct, hr, lactate, sex), 
                       10)

hbtot_final1 %>% 
  tidy %>% 
  knitr::kable() # AIC 86.9

```

status ~ Hb_tot + age_calc + bp_sys + bp_dias + hr + hct + lactate + sex
```{r hb_tot_model2}

# 2. status ~ Hb_tot + age_calc + bp_sys + bp_dias + hr + hct + lactate + sex
hbtot_final2 <- f_run_lr(df_model %>% 
                           dplyr::select(subject_id, status, cerebral_hb_tot, 
                                         age_calc, bp_sys, bp_dias, hct, hr, lactate, sex), 
                       10)

hbtot_final2 %>% 
  tidy %>% 
  knitr::kable() # AIC 85.6

```

status ~ Hb_tot + age_calc + bp_dias + hr + hct + lactate + sex
```{r hb_tot_model3}

# 3. status ~ Hb_tot + age_calc + bp_dias + hr + hct + lactate + sex
hbtot_final3 <- f_run_lr(df_model %>% 
                           dplyr::select(subject_id, status, cerebral_hb_tot, 
                                         age_calc, bp_dias, hct, hr, lactate, sex), 
                       10)

hbtot_final3 %>% 
  tidy %>% 
  knitr::kable() # AIC 83.52

```

status ~ Hb_tot + bp_dias + hr + hct + lactate + sex
```{r hb_tot_model4}

# 4. status ~ Hb_tot + bp_dias + hr + hct + lactate + sex
hbtot_final4 <- f_run_lr(df_model %>% 
                           dplyr::select(subject_id, status, cerebral_hb_tot, 
                                         bp_dias, hct, hr, lactate, sex), 
                       10)

hbtot_final4 %>% 
  tidy %>% 
  knitr::kable() # AIC 82.83

```

status ~ Hb_tot + hr + hct + lactate + sex
```{r hb_tot_model5}

# 5. status ~ Hb_tot + hr + hct + lactate + sex
hbtot_final5 <- f_run_lr(df_model %>% 
                           dplyr::select(subject_id, status, cerebral_hb_tot, 
                                         hct, hr, lactate, sex), 
                       10)

hbtot_final5 %>% 
  tidy %>% 
  knitr::kable() # AIC 81.36

```

status ~ Hb_tot + hct + lactate + sex
```{r hb_tot_model6}

# 6. status ~ Hb_tot + hct + lactate + sex

set.seed(20221118)

# resample
hbtot_boot <- bootstraps(df_model, strata = status, times = 5000)

# recipe
hbtot_rec <- recipe(status ~ cerebral_hb_tot + hct + lactate + sex + subject_id, data = df_model) %>% 
  step_impute_knn(all_predictors()) %>% 
  step_normalize(all_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_smote(status) %>% 
  update_role(subject_id, new_role = "id")

# create workflow
hbtot_wf <- workflow() %>% add_recipe(hbtot_rec)

# logistic regression model specifications
hbtot_glm_spec <- logistic_reg() %>% set_engine("glm")

# create model on 5000 different resamples
hbtot_glm_res <- hbtot_wf %>% 
  add_model(hbtot_glm_spec) %>%
  fit_resamples(
    resamples = hbtot_boot,
    metrics = metric_set(roc_auc, accuracy, sensitivity, specificity),
    control = tune::control_resamples(save_pred = TRUE, verbose = TRUE)
  )

# select best model to be used for evaluation
hbtot_glm_best <- hbtot_glm_res %>% select_best("roc_auc")

# finalize model
hbtot_final6 <- hbtot_wf %>% 
  add_model(hbtot_glm_spec) %>% 
  finalize_workflow(hbtot_glm_best) %>%
  fit(df_model) %>%
  extract_fit_parsnip()

hbtot_final6 %>% 
  tidy %>% 
  knitr::kable() # AIC 82.17, every term is significant

```

```{r assess_hb_tot_model}

df_CI <- confint(hbtot_final6$fit) %>% exp() %>% as_tibble(rownames = "term")

hbtot_final6 %>%
  tidy(exponentiate = TRUE) %>%
  filter(p.value < 0.05) %>% 
  left_join(df_CI, by = "term") %>%
  mutate(estimate = ifelse(estimate < 10^-2, scientific(estimate, digits = 3), round(estimate, 2)),
         conf.low = ifelse(`2.5 %` < 10^-2, scientific(`2.5 %`, digits = 3), round(`2.5 %`, 2)),
         conf.high = ifelse(`97.5 %` < 10^-2, scientific(`97.5 %`, digits = 3), round(`97.5 %`, 2))) %>%
  mutate(p.value = scientific(p.value, digits = 3),
         statistic = round(statistic, 2)) %>%
  unite("95% CI", c(conf.low, conf.high), sep = ", ") %>%
  dplyr::select(c(term, estimate, `95% CI`, statistic, p.value)) %>% 
  dplyr::rename("Term" = "term", "Odds ratio" = "estimate",
                "t-statistic" = "statistic", "p-value" = "p.value") %>%
  filter(Term != "(Intercept)") %>% 
  as.data.frame() %>%

  knitr::kable(align = rep('c', 5))

hbtot_glm_res %>% 
  collect_metrics() %>% 
  dplyr::select(.metric, mean, std_err) %>% 
  mutate(mean = mean*100, std_err = std_err*100) %>% 
  mutate(Metric = c("Accuracy", "ROC AUC", "Sensitivity", "Specificity")) %>% 
  dplyr::select(c("Metric", "mean", "std_err")) %>%

  dplyr::rename("Mean" = "mean", "Standard error" = "std_err") %>% 
  mutate_if(is.numeric, round, 2) %>% 
  as.data.frame() %>%

  knitr::kable()

```

```{r hb_tot_imp, fig.height = 4}

set.seed(123)
hbtot_imp <- hbtot_wf %>% 
  add_model(hbtot_glm_spec) %>% 
  fit(df_model) %>% 
  pull_workflow_fit() %>%

  vi(
    method = "permute", nsim = 10,
    train = juice(hbtot_rec %>% prep()),
    target = "status", metric = "auc", reference_class = "CM",
    pred_wrapper = kernlab::predict

  )

hbtot_imp %>% 
  clean_names %>% 
  filter(variable != "subject_id") %>% 
  mutate(variable = fct_reorder(variable, importance)) %>%

  ggplot(aes(x = importance, y = variable, color = variable)) +
  geom_errorbar(aes(xmin = importance - st_dev, xmax = importance + st_dev),
                alpha = 0.5, size = 1.3) +
  geom_point(size = 3) +
  theme(legend.position = "none",
        plot.title = element_text(size = 18),
        axis.title = element_text(size = 15),
        axis.text = element_text(size = 12),
        strip.text = element_text(size = 12),
        legend.title = element_text(size = 15),
        legend.text = element_text(size = 12)) +
  labs(y = NULL, x = "relative importance") +
  ggtitle("Variable importance in distinguishing CM vs UM")

```

### Cerebral Hb_tot_alpha model fitting

status ~ Hb_tot_alpha + age_calc + bp_sys + bp_dias + hr + hct + lactate + aa_arg + sex
```{r hb_tot_alpha_model1}

# 1. status ~ Hb_tot_alpha + age_calc + bp_sys + bp_dias + hr + hct + lactate + aa_arg + sex
hbtot_alpha_final1 <- f_run_lr(df_model %>% 
                           dplyr::select(subject_id, status, cerebral_hb_tot_alpha2, 
                                         aa_arg, age_calc, bp_sys, bp_dias, hct, hr, lactate, sex), 
                       10)

hbtot_alpha_final1 %>% 
  tidy %>% 
  knitr::kable() # AIC 64.6

```

status ~ Hb_tot_alpha + age_calc + bp_sys + bp_dias + hct + lactate + aa_arg + sex
```{r hb_tot_alpha_model2}

# 2. status ~ Hb_tot_alpha + age_calc + bp_sys + bp_dias + hct + lactate + aa_arg + sex
hbtot_alpha_final2 <- f_run_lr(df_model %>% 
                           dplyr::select(subject_id, status, cerebral_hb_tot_alpha2, 
                                         aa_arg, age_calc, bp_sys, bp_dias, hct, lactate, sex), 
                       10)

hbtot_alpha_final2 %>% 
  tidy %>% 
  knitr::kable() # AIC 67.99

```

status ~ Hb_tot_alpha + age_calc + bp_sys + bp_dias + hct + lactate + sex
```{r hb_tot_alpha_model3}

# 3. status ~ Hb_tot_alpha + age_calc + bp_sys + bp_dias + hct + lactate + sex
hbtot_alpha_final3 <- f_run_lr(df_model %>% 
                           dplyr::select(subject_id, status, cerebral_hb_tot_alpha2, 
                                         age_calc, bp_sys, bp_dias, hct, lactate, sex), 
                       10)

hbtot_alpha_final3 %>% 
  tidy %>% 
  knitr::kable() # AIC 61.78

```

status ~ Hb_tot_alpha + bp_sys + bp_dias + hct + lactate + sex
```{r hb_tot_alpha_model4}

# 4. status ~ Hb_tot_alpha + bp_sys + bp_dias + hct + lactate + sex
hbtot_alpha_final4 <- f_run_lr(df_model %>% 
                           dplyr::select(subject_id, status, cerebral_hb_tot_alpha2, 
                                         bp_sys, bp_dias, hct, lactate, sex), 
                       10)

hbtot_alpha_final4 %>% 
  tidy %>% 
  knitr::kable() # AIC 60.53

```

status ~ Hb_tot_alpha + bp_dias + hct + lactate + sex
```{r hb_tot_alpha_model5}

# 5. status ~ Hb_tot_alpha + bp_dias + hct + lactate + sex
hbtot_alpha_final5 <- f_run_lr(df_model %>% 
                           dplyr::select(subject_id, status, cerebral_hb_tot_alpha2, 
                                         bp_dias, hct, lactate, sex), 
                       10)

hbtot_alpha_final5 %>% 
  tidy %>% 
  knitr::kable() # AIC 61.65

```

status ~ Hb_tot_alpha + hct + lactate + sex
```{r hb_tot_alpha_model6}

# 6. status ~ Hb_tot_alpha + hct + lactate + sex
hbtot_alpha_final6 <- f_run_lr(df_model %>% 
                           dplyr::select(subject_id, status, cerebral_hb_tot_alpha2, 
                                         hct, lactate, sex), 
                       10)

hbtot_alpha_final6 %>% 
  tidy %>% 
  knitr::kable() # AIC 67.5

```

status ~ Hb_tot_alpha + hct + sex
```{r hb_tot_alpha_model7}

# 7. status ~ Hb_tot_alpha + hct + sex

set.seed(20221118)

# resample
hbtot_alpha_boot <- bootstraps(df_model, strata = status, times = 5000)

# recipe
hbtot_alpha_rec <- recipe(status ~ cerebral_hb_tot_alpha2 + hct + sex + subject_id, data = df_model) %>% 
  step_impute_knn(all_predictors()) %>% 
  step_normalize(all_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_smote(status) %>% 
  update_role(subject_id, new_role = "id")

# create workflow
hbtot_alpha_wf <- workflow() %>% add_recipe(hbtot_alpha_rec)

# logistic regression model specifications
hbtot_alpha_glm_spec <- logistic_reg() %>% set_engine("glm")

# create model on 5000 different resamples
hbtot_alpha_glm_res <- hbtot_alpha_wf %>% 
  add_model(hbtot_alpha_glm_spec) %>%
  fit_resamples(
    resamples = hbtot_alpha_boot,
    metrics = metric_set(roc_auc, accuracy, sensitivity, specificity),
    control = tune::control_resamples(save_pred = TRUE, verbose = TRUE)
  )

# select best model to be used for evaluation
hbtot_alpha_glm_best <- hbtot_alpha_glm_res %>% select_best("roc_auc")

# finalize model
hbtot_alpha_final7 <- hbtot_alpha_wf %>% 
  add_model(hbtot_alpha_glm_spec) %>% 
  finalize_workflow(hbtot_alpha_glm_best) %>%
  fit(df_model) %>%
  extract_fit_parsnip()

hbtot_alpha_final7 %>% 
  tidy %>% 
  knitr::kable() # AIC 65.23

```

```{r assess_hb_tot_alpha_model}

df_CI <- confint(hbtot_alpha_final7$fit) %>% exp() %>% as_tibble(rownames = "term")

hbtot_alpha_final7 %>%
  tidy(exponentiate = TRUE) %>%
  filter(p.value < 0.05) %>% 
  left_join(df_CI, by = "term") %>%
  mutate(estimate = ifelse(estimate < 10^-2, scientific(estimate, digits = 3), round(estimate, 2)),
         conf.low = ifelse(`2.5 %` < 10^-2, scientific(`2.5 %`, digits = 3), round(`2.5 %`, 2)),
         conf.high = ifelse(`97.5 %` < 10^-2, scientific(`97.5 %`, digits = 3), round(`97.5 %`, 2))) %>%
  mutate(p.value = scientific(p.value, digits = 3),
         statistic = round(statistic, 2)) %>%
  unite("95% CI", c(conf.low, conf.high), sep = ", ") %>%
  dplyr::select(c(term, estimate, `95% CI`, statistic, p.value)) %>% 
  dplyr::rename("Term" = "term", "Odds ratio" = "estimate",
                "t-statistic" = "statistic", "p-value" = "p.value") %>%
  filter(Term != "(Intercept)") %>% 
  as.data.frame() %>%

  knitr::kable(align = rep('c', 5))

hbtot_alpha_glm_res %>% 
  collect_metrics() %>% 
  dplyr::select(.metric, mean, std_err) %>% 
  mutate(mean = mean*100, std_err = std_err*100) %>% 
  mutate(Metric = c("Accuracy", "ROC AUC", "Sensitivity", "Specificity")) %>% 
  dplyr::select(c("Metric", "mean", "std_err")) %>%

  dplyr::rename("Mean" = "mean", "Standard error" = "std_err") %>% 
  mutate_if(is.numeric, round, 2) %>% 
  as.data.frame() %>%

  knitr::kable()

```

```{r hb_tot_alpha_imp, fig.height = 4}

set.seed(123)
hbtot_alpha_imp <- hbtot_alpha_wf %>% 
  add_model(hbtot_alpha_glm_spec) %>% 
  fit(df_model) %>% 
  pull_workflow_fit() %>%

  vi(
    method = "permute", nsim = 10,
    train = juice(hbtot_alpha_rec %>% prep()),
    target = "status", metric = "auc", reference_class = "CM",
    pred_wrapper = kernlab::predict

  )

hbtot_alpha_imp %>% 
  clean_names %>% 
  filter(variable != "subject_id") %>% 
  mutate(variable = fct_reorder(variable, importance)) %>%

  ggplot(aes(x = importance, y = variable, color = variable)) +
  geom_errorbar(aes(xmin = importance - st_dev, xmax = importance + st_dev),
                alpha = 0.5, size = 1.3) +
  geom_point(size = 3) +
  theme(legend.position = "none",
        plot.title = element_text(size = 18),
        axis.title = element_text(size = 15),
        axis.text = element_text(size = 12),
        strip.text = element_text(size = 12),
        legend.title = element_text(size = 15),
        legend.text = element_text(size = 12)) +
  labs(y = NULL, x = "relative importance") +
  ggtitle("Variable importance in distinguishing CM vs UM")

```

### Distributions of NIRS variables by sex

```{r sex_associations}

df_hbox %>% 
  dplyr::select(subject_id, status, sex, cerebral_hb_tot, cerebral_hb_tot_alpha2) %>% 
  pivot_longer(4:5, names_to = "nirs_variable", values_to = "value") %>%

  ggplot(aes(x = status, y = value)) +
  geom_violin(aes(color = status)) +
  geom_jitter(position = position_jitter(0.15), shape = 1, size = 2) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = status), size = 0.2, width = 0.5) +
  facet_grid(nirs_variable ~ sex, scales = "free")

# check for significant differences between M-F for each variable
df_hbox %>% 
  dplyr::select(subject_id, status, sex, cerebral_hb_tot, cerebral_hb_tot_alpha2) %>% 
  pivot_longer(4:5, names_to = "nirs_variable", values_to = "value") %>%

  group_by(status, nirs_variable) %>% 
  nest() %>%

  mutate(wilcox = map(.x = data,
                      .f = ~ wilcox.test(.x %>% filter(sex == "Female") %>% pull(value),
                                         .x %>% filter(sex == "Male") %>% pull(value)))) %>% 
  mutate(p_val = map(.x = wilcox, .f = ~.x$p.value)) %>% 
  unnest(cols = c(p_val)) %>%

  knitr::kable()

```

**no sex differences in distributions of NIRS variables in each patient group**